---
title: "Graph: La evolución del despligue militar de EE.UU."
description:
  "This project will explain the process of replicating the graphic La evolución del despliegue militar de EE.UU. created by El Orden Mundial."
categories: "2023"
author: Clara Fernanda Martín Villaescusa
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    self_contained: false
    toc: true 
---

# REPLICATION

## DATA BASE

```{r}
library(tidyverse)
library(scales)
library(here)
library(haven)
library(ggtext)
library(troopdata)
library(glue)
library(ggtext)
library(ggplotify)
library(treemapify)
library(ggalluvial)
library(ggpmisc)
library(ggrepel)
library(patchwork)
library(scico)
library(ggbeeswarm)
library(peacesciencer)
library(kableExtra)
library(gridExtra)
library(ggpmisc)
library(dplyr)
#install.packages("Hmisc")
library(Hmisc)

#install.packages("showtext")
library(showtext)



troopdata <- troopdata::troopdata |>  
  filter(ccode != 2 & year >= 1950) |> 
  mutate(ccode = ifelse(ccode == 260, 255, ccode),
         ccode = ifelse(countryname == "Republic of Vietnam", 816, ccode),
         iso3c = ifelse(countryname == "Republic of Vietnam", "VNM", iso3c),
         region = countrycode::countrycode(iso3c, "iso3c", "region"),
         region = case_when(
           ccode == 1003 ~ "Afloat",
           ccode == 70 ~ "North America",
           region == "Latin America & Caribbean" ~ "Central & South America & Caribbean", 
           TRUE ~ region), 
         region = recode(region, "Central & South America & Caribbean" = "Latinoamérica y Caribe", "Europe & Central Asia" = "Europa y Asia Central", "Middle East & North Africa" = "Oriente Próximo y Magreb", "Sub-Saharan Africa" = "África Subsahariana", "South Asia" = "Asia Meridional", "East Asia & Pacific" = "Asia Oriental y Oceanía")) 
```

### Troopdata without highlighted countries

```{r}
eliminar <- c("Germany", "West Germany", "German Democratic Republic", "Vietnam", "Japan", "South Korea", "Iraq")

troopdata_sin <- troopdata::troopdata |>  
  filter(ccode != 2 & year >= 1950) |> 
  mutate(
         region = countrycode::countrycode(iso3c, "iso3c", "region"),
         region = case_when(
           ccode == 1003 ~ "Afloat",
           ccode == 70 ~ "North America",
           region == "Latin America & Caribbean" ~ "Central & South America & Caribbean", 
           TRUE ~ region), 
         region = recode(region, "Central & South America & Caribbean" = "Latinoamérica y Caribe", "Europe & Central Asia" = "Europa y Asia Central", "Middle East & North Africa" = "Oriente Próximo y Magreb", "Sub-Saharan Africa" = "África Subsahariana", "South Asia" = "Asia Meridional", "East Asia & Pacific" = "Asia Oriental y Oceanía")) |> 
  filter(!(countryname %in% eliminar))
```

## TIDY DATA

### REGION DATA

I do the region data with troopdata without the highlighted countries so that they are not counted twice.

```{r}
regional.data <- troopdata_sin |> 
  group_by(region, year) |>  
  dplyr::summarise(troopsum = sum(troops, na.rm = TRUE)) |>  
  filter(region != 'Afloat') |> 
  filter(region != 'North America') 


regional.data <- subset(regional.data, year <= 2020)

regional.data$region <- factor(regional.data$region , levels= c("Europa y Asia Central", "África Subsahariana", "Asia Oriental y Oceanía", "Latinoamérica y Caribe", "Asia Meridional", "Oriente Próximo y Magreb"))
```

### COUNTRY DATA

here are only the highlighted countries. Germany is recategorised because there are several germanies in the database.

```{r}
country.data <- troopdata |>
  filter((grepl("Vietnam", countryname)) | grepl('Germany', countryname) | grepl ("South Korea", countryname) | grepl ("Japan", countryname) | grepl ("Iraq", countryname)) |>
  mutate(countryname = ifelse(countryname == "Republic of Vietnam", "Vietnam", countryname),
            countryname = ifelse(countryname == "West Germany", "Germany", countryname),
         countryname = ifelse(countryname == "German Democratic Republic", "Germany", countryname)
         )|>
  group_by(countryname, year) |>  
  dplyr::summarise(troopsum = sum(troops, na.rm = TRUE)) |>
  mutate(countryname = factor(countryname, 
                              levels = c("Iraq", "Japan", 
                                         "South Korea", "Vietnam",
                                         "Germany", 
                                         setdiff(unique(countryname),
                                                 c("Iraq", "Japan",
                                                   "South Korea",
                                                   "Vietnam",
                                                   "Germany")))))

country.data$countryname <- factor(country.data$countryname , levels=rev(c("Germany", "Vietnam", "Japan", "South Korea", "Iraq")))
```

### COMBINATION OF REGION AND COUNTRY

combination of the two data frames created above

```{r}
combined_data <- bind_rows(
  mutate(regional.data, data_type = "region"),
  mutate(country.data, data_type = "country"))

combined_data <- bind_rows(regional.data, country.data)

combined_data <- combined_data |> 
  pivot_longer(cols = c(region, countryname), names_to = "tipo", values_to = "region_country")
```

```{r}
#sort categories 
combined_data$region_country <- factor(combined_data$region_country , levels = c("Latinoamérica y Caribe", "África Subsahariana","Iraq", "Oriente Próximo y Magreb", "Asia Meridional","South Korea","Japan","Vietnam","Asia Oriental y Oceanía", "Germany","Europa y Asia Central"))
```

## GGPLOT

### THEME

cosas que faltan: Poner mas ticks en el eje X y ponerlos por dentro.

```{r}
sysfonts::font_add_google("Outfit", family="outfit")
showtext::showtext_auto()

theme_ok <- theme(text = element_text(family="outfit", size = 11),
  plot.title = element_text(face = "bold", size = 18),
  plot.subtitle = element_text(face = "bold", color = "gray30" ,size = 12),
  plot.title.position = "plot",
  plot.caption = ggtext::element_markdown(size = rel(0.7), lineheight = 0.9, hjust = 0, margin = margin(0, 0, 0, 0)),
  plot.caption.position = "plot",
  strip.background = element_rect(fill = "gray80", color = "black"),
  panel.background = element_rect(fill = "white"),
  strip.text = element_text(color = "black", face = "bold"),
  panel.grid.major.y = element_line(color = "lightgrey"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.line = element_line(color = "black",size = 1),
  axis.title.y = element_text(angle = 90, hjust = 0.5, color = "gray40"),
  axis.text.x = element_text(angle = 45, hjust = 1),
  legend.position = "top",
  legend.direction = "horizontal",
  legend.title = element_blank(),  # Remove legend title
  legend.text = element_text(hjust = 0, face = "plain"),
  legend.background = element_blank(),  # Remove legend background
  legend.box = "none",  # Remove legend box
  legend.justification = "left",  
  legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
  legend.key = element_blank(),
  legend.key.size = unit(0.5, "cm"),
  legend.key.width = unit(0.8, "cm"),
  legend.key.height = unit(0.5, "cm"))

```

```{r}
colors <- c("Europa y Asia Central" = "#676B90", "Asia Oriental y Oceanía" = "#DE5A3C", "Asia Meridional" = "#70B3B5", "Oriente Próximo y Magreb" = "#8BA629", "África Subsahariana" = "#9E65A6", "Latinoamérica y Caribe" = "#F2BE5C", "Iraq" = "#AACE7B", "Japan" = "#DEA281", "South Korea" = "#E5C5B1", "Vietnam" = "#DF8756", "Germany" = "#7185B5")

colors2 <- c("Europa y Asia Central" = "#676B90", "Asia Oriental y Oceanía" = "#DE5A3C", "Asia Meridional" = "#70B3B5", "Oriente Próximo y Magreb" = "#8BA629", "África Subsahariana" = "#9E65A6", "Latinoamérica y Caribe" = "#F2BE5C", "Iraq" = "#8BA629", "Japan" = "#DEA281", "South Korea" = "#DE5A3C", "Vietnam" = "#DF8756", "Germany" = "#676B90")
```

### REGION GRAPH

```{r}
regional.time <- ggplot(regional.data, aes(x = year, y = troopsum, fill = region)) +
  geom_area() +
  scale_x_continuous(
    limits = c(1950, 2020),
    breaks = seq(1950, 2020, 10),
    expand = expansion(add = c(1.5, 0))) +
  scale_y_continuous(labels = scales::number_format(big.mark = ".", decimal.mark = ",", accuracy = 1), 
                     limits = c(0, 1200000),
                     breaks = seq(0, 1200000, 200000),
                     expand = expansion(0)) +
  scale_fill_manual(values = c("Europa y Asia Central" = "#676B90", "Asia Oriental y Oceanía" = "#D94A3D", "Asia Meridional" = "#70B3B5", "Oriente Próximo y Magreb" = "#8BA629", "África Subsahariana" = "#9E65A6", "Latinoamérica y Caribe" = "#F2BE5C")) +
  theme_ok +
  labs(x = "",
       y = "Soldados",
       color = "Region",
       linetype = "Region",
       title = "La evolución del despligue militar de EE.UU.",
       subtitle = "Presencia de tropas por regiones y países más destacados (1950-2020)") 

print(regional.time)
```

### COUNTRY GRAPH

```{r}
country.time <- ggplot(country.data, aes(x = year, y = troopsum, fill = countryname)) +
  geom_area() +
  scale_x_continuous(limits = c(1950, 2020), breaks = seq(1950, 2020, 10), expand = expansion(0.02)) +
  scale_y_continuous(labels = scales::number_format(big.mark = ".", decimal.mark = ",", accuracy = 1), 
                     limits = c(0, 1200000),
                     breaks = seq(0, 1200000, 200000),
                     expand = expansion(0))  +
  theme_ok + 
  labs(x = "",
       y = "Soldados",
       color = "Region",
       linetype = "Region",
       title = "La evolución del despligue militar de EE.UU.",
       subtitle = "Presencia de tropas por regiones y países más destacados (1950-2020)")  

print(country.time)
```

### COMBINED GRAPH ✅

```{r}
combinado.time <- ggplot() +
  geom_area(data = combined_data |>  filter(!is.na(region_country)), 
            mapping = aes(x = year, y = troopsum, fill = region_country, color = region_country), 
            position = "stack", size = 1.3) +  
  scale_x_continuous(
    limits = c(1950, 2020),
    breaks = seq(1950, 2020, 5),
    expand = expansion(add = c(0.5, 0)))+
    #minor_breaks = seq(1950, 2020, by = 5)) +
  scale_y_continuous(labels = scales::number_format(big.mark = ".", decimal.mark = ",", accuracy = 1),
                     limits = c(0, 1200000),
                     breaks = seq(0, 1200000, 200000),
                     expand = expansion(0)) +
   scale_fill_manual(name = "Leyenda", values = colors, breaks = c("Europa y Asia Central", "África Subsahariana", "Asia Oriental y Oceanía", "Latinoamérica y Caribe", "Asia Meridional", "Oriente Próximo y Magreb")) +
  scale_color_manual(values = colors2, breaks = unique(combined_data$region_country), guide = FALSE) +
  theme_ok +
  labs(
    x = "",
    y = "Soldados",
    color = "Data Type",
    fill = "Data Type",
    title = "La evolución del despliegue militar de EE.UU.",
    subtitle = "Presencia de tropas por regiones y países más destacados (1950-2020)", 
    caption = paste("**Gráfico:** 
                    Álvaro Merino (2021) 
                    **Fuente:** 
                    Michael A. Allen, Michael E. Flynn y Carla Martínez Machain, Us global military deployments, 1950-2020 (2021)"))

combinado.time
```

### Annotations

(falta cambiar el tamaño de la cabeza de las fechas, no se hacerlo)

```{r}
combinado.timeA <- combinado.time + annotate("text",
  label = "Alemania",
  x = 1980, y = 150000, hjust = 0.3, vjust = 0, size = 4, color = "white", family="outfit") + 
  annotate("text",
  label = "Vietnam",
  x = 1968, y = 450000, hjust = 0.4, vjust = 0, size = 4, color = "white", family="outfit") +
  annotate("text",
  label = "Irak",
  x = 2006, y = 300000, hjust = 0.6, vjust = 0, size = 4, color = "white", family="outfit") +
  annotate("text",
  label = "Japón",
  x = 1955, y = 520000, hjust = 0.5, vjust = 0, size = 4, color = "white", family="outfit") +
   annotate("text",
  label = "11S",
  x = 2001, y = 440000, hjust = 0.5, vjust = 0, size = 4, color = "gray30", family="outfit") +
  annotate("text",
  label = "Colapso de la URSS",
  x = 1992, y = 540000, hjust = 0.09, vjust = 0, size = 4, color = "gray30", family="outfit") +
  annotate("text",
  label = "Caída del Muro \nde Berlín",
  x = 1990, y = 700000, hjust = 0.5, vjust = -0.4, size = 4, color = "gray30", family="outfit", lineheight = 0.4) +
  annotate("segment", x = 1990, xend = 1990, y = 0, yend = 700000, color = "gray50", linetype = "aa") +
  annotate("segment", x = 1992, xend = 1992, y = 0, yend = 520000, color = "gray50", linetype = "aa") +
  annotate("segment", x = 2001, xend = 2001, y = 0, yend = 420000, color = "gray50", linetype = "aa") +
  annotate("segment", x = 2012, xend = 2014, y = 200000, yend = 330000, color = "gray30", linetype = "solid") +
  annotate("text",
  label = "Afganistán",
  x = 2014, y = 330000, hjust = 0.3, vjust = -0.5, size = 4, color = "gray30", family="outfit") +
  annotate("segment", x = 1953, xend = 1957, y = 700000, yend = 840000, color = "gray30", linetype = "solid") +
  annotate("text",
  label = "Corea \ndel Sur",
  x = 1958, y = 800000, hjust = 0.2, vjust = -0.2, size = 4, color = "gray30", family="outfit", lineheight = 0.4) +
  annotate("curve", x = 2010, y = 210000 , xend = 2010, yend = 640000, color = "gray30", curvature = 0.06, arrow = arrow(length = unit(0.01, "npc"), angle = 20, type = "closed")) +
  annotate("curve", x = 2006, y = 350000 , xend = 2007, yend = 640000, color = "gray30", curvature = -0.1, arrow = arrow(length = unit(0.01, "npc"), angle = 20, type = "closed")) +
  annotate("text",
  label = "Guerra contra \nel terror",
  x = 2009, y = 660000, hjust = 0.5, vjust = 0, size = 4, color = "gray30", family="outfit", lineheight = 0.4) +
  annotate("curve", x = 1968, y = 850000 , xend = 1975, yend = 920000, color = "gray30", curvature = -0.2, arrow = arrow(length = unit(0.01, "npc"), angle = 20, type = "closed")) +
  annotate("text",
  label = "Guerra de Vietnam",
  x = 1975, y = 920000, hjust = -0.05, vjust = 0.3, size = 4, color = "gray30", family="outfit") +
  annotate("curve", x = 1953, y = 880000 , xend = 1957, yend = 1100300, color = "gray30", curvature = -0.3, arrow = arrow(length = unit(0.01, "npc"), angle = 20, type = "closed")) +
  annotate("text",
  label = "Guerra de Corea",
  x = 1957, y = 1100300, hjust = -0.05, vjust = -0.3, size = 4, color = "gray30", family="outfit") 

combinado.timeA

```


# ALTERNATIVE

## Data Base.

```{r}
troopdata1 <- troopdata::troopdata |>  
  filter(ccode != 2 & year >= 1950) |> 
  mutate(ccode = ifelse(ccode == 260, 255, ccode),
         ccode = ifelse(countryname == "Republic of Vietnam", 816, ccode),
         iso3c = ifelse(countryname == "Republic of Vietnam", "VNM", iso3c),
         region = countrycode::countrycode(iso3c, "iso3c", "region"),
         region = case_when(
           ccode == 1003 ~ "Afloat",
           ccode == 70 ~ "North America",
           region == "Latin America & Caribbean" ~ "Central & South America & Caribbean", 
           TRUE ~ region), 
         region = recode(region, "Central & South America & Caribbean" = "Latinoamérica y Caribe", "Europe & Central Asia" = "Europa y Asia Central", "Middle East & North Africa" = "Oriente Próximo y Magreb", "Sub-Saharan Africa" = "África Subsahariana", "South Asia" = "Asia Meridional", "East Asia & Pacific" = "Asia Oriental y Oceanía")) 
```

## TidyData: Region

```{r}
regional.data1 <- troopdata1 |> 
  group_by(region, year) |> 
  dplyr::summarise(troopsum = sum(troops, na.rm = TRUE)) |> 
  filter(region != 'Afloat') |>
  filter(region != 'North America')
regional.data1$region <- factor(regional.data1$region , levels= c("Europa y Asia Central", "África Subsahariana", "Asia Oriental y Oceanía", "Latinoamérica y Caribe", "Asia Meridional", "Oriente Próximo y Magreb"))
```

## Theme Alternative.

```{r}
sysfonts::font_add_google("Merriweather", family="Merriweather")
showtext::showtext_auto()

theme_ok1 <- theme(text = element_text(family="Merriweather", size = 16),
  plot.title = element_text(face = "bold", size = 23),
  plot.subtitle = element_text(face = "bold", color = "gray30" ,size = 18),
  plot.title.position = "plot",
  plot.caption = ggtext::element_markdown(size = rel(0.7), lineheight = 0.9, hjust = 0, margin = margin(0, 0, 0, 0)),
  plot.caption.position = "plot",
  strip.background = element_rect(fill = "gray80", color = "black"),
  panel.background = element_rect(fill = "white"),
  strip.text = element_text(color = "black", face = "bold"),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.line = element_line(color = "black",size = 0.6),
  axis.title.y = element_text(angle = 90, hjust = 0.5),
  axis.text.x = element_text(angle = 45, hjust = 1),
  legend.position = "top",
  legend.direction = "horizontal",
  legend.title = element_blank(),  
  legend.text = element_text(hjust = 0),
  legend.background = element_blank(),  
  legend.box = "none",  
  legend.justification = c(-0.2, 0),  
  legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
  legend.key = element_blank(),
  legend.key.size = unit(0.5, "cm"))
```

## Graph Alternative.

```{r}
regional.time1 <- ggplot() +
  geom_line(data = regional.data1, mapping = aes(x = year, y = troopsum, color = region, linetype = region), linetype = "solid", size = 0.5) +  
  scale_x_continuous(limits = c(1950, 2020), breaks = seq(1950, 2020, 5), expand = expansion(add = c(1.5, 0))) +
  scale_y_continuous(labels = scales::number_format(big.mark = ".", decimal.mark = ",", accuracy = 1), breaks = seq(0, 1000000, 100000)) +
  scale_color_manual(name = NULL, values = c("Europa y Asia Central" = "#676B90", "Asia Oriental y Oceanía" = "#DE5A3C", "Asia Meridional" = "#70B3B5", "Oriente Próximo y Magreb" = "#8BA629", "África Subsahariana" = "#9E65A6", "Latinoamérica y Caribe" = "#F2BE5C")) +
  theme_ok1  +
  labs(
    x = "",
    y = "Soldados",
    title = "La evolución del despliegue militar de EE.UU.",
    subtitle = "Presencia de tropas por regiones (1950-2020)", 
    caption = paste("Fuente: Michael A. Allen, Michael E. Flynn y Carla Martínez Machain, Us global military deployments, 1950-2020 (2021)")
  )

regional.time1 <- regional.time1 + annotate("text",
  label = "11S",
  x = 2001, y = 230000, hjust = 0.5, vjust = 0.6, size = 5, color = "gray30", family="Merriweather") +
  annotate("text",
  label = "Colapso de la URSS",
  x = 1992, y = 360000, hjust = 0.05, vjust = 0.7, size = 5, color = "gray30", family="Merriweather") +
  annotate("text",
  label = "Caída del Muro de Berlín",
  x = 1990, y = 470000, hjust = 0.5, vjust = 0, size = 5, color = "gray30", family="Merriweather") +
  annotate("segment", x = 1990, xend = 1990, y = 0, yend = 450000, color = "gray50", linetype = "aa", size = 0.2) +
  annotate("segment", x = 1992, xend = 1992, y = 0, yend = 340000, color = "gray50", linetype = "aa", size = 0.2) +
  annotate("segment", x = 2001, xend = 2001, y = 0, yend = 200000, color = "gray50", linetype = "aa", size = 0.2) +
  annotate("text",
  label = "Guerra contra \nel terror",
  x = 2013, y = 230000, hjust = 0.5, vjust = 0, size = 5, color = "gray30", lineheight = 0.4, family="Merriweather") +
  annotate("curve", x = 2012, y = 50000 , xend = 2013, yend = 200000, color = "#70B3B5", curvature = 0.06, arrow = arrow(length = unit(0.01, "npc"), angle = 20, type = "closed")) +
  annotate("curve", x = 2007, y = 150000 , xend = 2010, yend = 240000, color = "#8BA629", curvature = -0.1, arrow = arrow(length = unit(0.01, "npc"), angle = 20, type = "closed")) +
  annotate("curve", x = 1968, y = 650000 , xend = 1971, yend = 710000, color = "#DE5A3C", curvature = -0.1, arrow = arrow(length = unit(0.01, "npc"), angle = 20, type = "closed")) +
  annotate("text",
  label = "Guerra de Vietnam",
  x = 1972, y = 720000, hjust = 0, vjust = 0, size = 5, color = "gray30", family="Merriweather") +
   annotate("curve", x = 1953, y = 490000 , xend = 1956, yend = 520000, color = "#DE5A3C", curvature = -0.1, arrow = arrow(length = unit(0.01, "npc"), angle = 20, type = "closed")) +
  annotate("text",
  label = "Ocupación \nde Japón",
  x = 1957, y = 620000, hjust = 0.05, vjust = 2, size = 5, color = "gray30", lineheight = 0.4,family="Merriweather")

regional.time1
```


