---
title: "Track National Unemployment"
description: |
  National Unemployment, Job Gains and Job Losses.
categories: "2025"
author: Miriam Lim Leal
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    self_contained: false
    toc: true
---
## Libraries

```{r}
library(tidyverse)
library(cowplot)
library(colorspace)
```

## 1.Import data

```{r}
payrol_unemployment <- read.csv(file = "./monthly.csv")
```

## 2.Add names to variables and convert "date" variable into date

```{r}
new_payrol_unemployement <- 
  payrol_unemployment |> 
  rename(date = "observation_date",
         transportation_and_warehousing = "CES4300000001",
         utilities = "CES4422000001",
         education = "CES6561000001",
         health_care_and_social_assistance = "CES6562000001",
         arts_entertainment_and_rec = "CES7071000001",
         accommodation = "CES7072100001",
         restaurants_and_bars = "CES7072200001",
         federal_government = "CES9091000001",
         state_government = "CES9092000001",
         local_government = "CES9093000001",
         durable_goods_manufacturing = "DMANEMP",
         nondurable_goods_manufacturing = "NDMANEMP",
         construction = "USCONS",
         financial_activities = "USFIRE",
         information = "USINFO",
         mining_and_logging = "USMINE",
         professional_and_business_services = "USPBS",
         other_services = "USSERV",
         retail_trade = "USTRADE",
         wholesale_trade = "USWTRADE"
         ) |> 
  mutate(date = as.Date(date, format = "%Y-%m-%d"))
```

## 3.Calculate month-over-month % change

```{r}
percentage_payroles <- new_payrol_unemployement |> 
  arrange(date) |> 
  mutate(across(
    c(transportation_and_warehousing, 
      utilities,
      education,
      health_care_and_social_assistance,
      arts_entertainment_and_rec,
      accommodation,
      restaurants_and_bars,
      federal_government,
      state_government,
      local_government,
      durable_goods_manufacturing,
      nondurable_goods_manufacturing,
      construction,
      financial_activities,
      information,
      mining_and_logging,
      professional_and_business_services,
      other_services,
      retail_trade,
      wholesale_trade),
    ~ (. - lag(.)) / lag(.) * 100
  ))
```

order data from january-feb to check

```{r}
percentage_payroles |>
  filter(date == "2008-02-01") |>
  select(-date) |>  # temporarily remove date
  pivot_longer(everything(), names_to = "industry", values_to = "pct_change") |>
  arrange(desc(pct_change)) |>
  select(industry, pct_change)
```

Data is slightly different than original because the BLS(Bureau of Labor Statistics) constantly revises employment numbers after they're first released, so data downloaded at different times will have slightly different values.

## 4. Split data into 2: stacked data and data_cero (values between -0.01 and 0.01)

### 4.1 plot_data (only values that are not between -0.01 and 0.01)

Reorganise data, dates as columns rest of info as rows

```{r}
plot_data <- percentage_payroles |> 
  pivot_longer(
    cols = -date,
    names_to = "sector",
    values_to = "percentage_change"
  ) |> 
  filter(date >= "2008-02-01" & date <= "2017-03-01") |> 
  filter(percentage_change <= -0.01 | percentage_change >= 0.01)
```

### 4.2 data_cero (only values between -0.01 and 0.01)

Reorganise data, dates as columns rest of info as rows

```{r}
data_zero <- percentage_payroles |> 
  pivot_longer(
    cols = -date,
    names_to = "sector",
    values_to = "percentage_change"
  ) |> 
  filter(date >= "2008-02-01" & date <= "2017-03-01") |> 
  filter(percentage_change > -0.01 & percentage_change < 0.01)
```

## 5.Sort and stack data

### 5.1 plot_data

```{r}
staked_data <- plot_data |> 
  group_by(date) |>  
  arrange(date, desc(percentage_change)) |> 
  mutate(is_positive = percentage_change > 0) |> 
  group_by(date, is_positive) |> 
    mutate(order = ifelse(is_positive,
                          rev(row_number()) - 0.5,
                          -row_number() + 0.5)
    ) |> 
  ungroup() |> 
  mutate(pc_bin = cut(   #create intervals with breaks saved in pc_bin
    percentage_change,
    breaks = c(-Inf, -1, -0.75, -0.5, -0.25, 0, 0.25, 0.5, 0.75, Inf),
    include.lowest = TRUE,
    labels = c("(-Inf,-1]", "(-1,-0.75]", "(-0.75,-0.5]", 
               "(-0.5,-0.25]", "(-0.25,0]", "(0,0.25]", 
               "(0.25,0.5]", "(0.5,0.75]", "(0.75,Inf]")
  ))

```

### 5.2 data_cero

```{r}
zero_stacked <- data_zero |> 
  group_by(date) |> 
  arrange(date, percentage_change) |> 
  mutate(stacked_position = row_number()) |> 
  ungroup()
  
```

## 6. Join data

```{r}
combined_data <- bind_rows(
  staked_data |> 
    mutate(category = "main") |> 
    select(date, order, pc_bin, category),
  
  zero_stacked |> 
    mutate(
      category = "unch",
      pc_bin = "unchanged"
    ) |> 
    mutate(order = -22 + (stacked_position - 1)) |>   # Separate mutate after
    select(date, order, pc_bin, category))
```

## 7. PLOTTING DATA

Create two tile plots joined

```{r}
p1 <- ggplot(combined_data, aes(x = date, y = order)) +
  annotate("rect", 
           xmin = as.Date(-Inf), 
           xmax = as.Date("2009-06-30"), 
           ymin = -Inf, 
           ymax = 20, 
           fill = "grey90", 
           alpha = 0.5) +
  annotate("text",
           x = as.Date("2008-02-01"),
           y = 16,
           label = "RECESSION",
           hjust = 0,
           vjust = 0.5,
           size = 2.5,
           color = "grey30") +
  
  geom_tile(data = subset(combined_data, 
                          category == "main"),
            aes(fill = pc_bin),
            width = 25, 
            height = 1, 
            colour = "white", 
            linewidth = 0.1) +
  
  geom_tile(data = subset(combined_data, 
                          category == "unch"),
            width = 25, 
            height = 1, 
            fill = "#999999", 
            colour = "white", 
            linewidth = 0.1) +
  
  scale_fill_manual(values = c(
    "(-Inf,-1]" = "#ce472e",
    "(-1,-0.75]" = "#f05336",
    "(-0.75,-0.5]" = "#faa224",
    "(-0.5,-0.25]" = "#ffd73e",
    "(-0.25,0]" = "#efe3be",
    "(0,0.25]" = "#c6e3bb",
    "(0.25,0.5]" = "#a3d393",
    "(0.5,0.75]" = "#64bc52",
    "(0.75,Inf]" = "#0099dc"
  )) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", position = "top", expand = c(0, 0)) +
  scale_y_continuous(breaks = c(20, 15, 10, 5, 0, -5, -10, -15, -20, -22),
                     labels = c("20", "15", "10", "5", "0", "-5", "-10", "-15", "-20", "Unch."),
                     limits = c(-23, 21),
                     expand = c(0, 0)) +
  coord_fixed(ratio = 25) +
  theme_minimal() +
  theme(
    axis.text.x.top = element_text(size = 6, angle = 0, hjust = 0.5, vjust = -1),
    axis.text.y = element_text(size = 6, hjust = 1, vjust = 0.5),
    axis.title.y = element_text(size = 5, angle = 90, vjust = 0.5),
    
    axis.ticks.x.top = element_line(color = "black", linewidth = 0.5),
    axis.ticks.length.x.top = unit(0.1, "cm"),
    
    axis.ticks.y = element_line(color = "black", linewidth = 0.5),
    axis.ticks.length.y = unit(0.1, "cm"),
    
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    
    plot.margin = margin(10, 10, 10, 10),
    legend.position = "none"
  ) +
  labs(x = NULL,
       y = "< SECTORS FALLING                SECTORS RISING > ",
       fill = NULL)
p1
```

Create a data for the legend of the colour scale

```{r}
legend_data <- data.frame(
  x = 1:9,
  y = 1,
  color_bin = c("(-Inf,-1]", "(-1,-0.75]", "(-0.75,-0.5]", "(-0.5,-0.25]", 
                "(-0.25,0]", "(0,0.25]", "(0.25,0.5]", "(0.5,0.75]", "(0.75,Inf]")
)
```

Create the plot

```{r}
p2 <- ggplot(legend_data, aes(x = x, y = y, fill = color_bin)) +
  geom_tile(width = 1, height = 1, color = NA) +
  scale_fill_manual(values = c(
    "(-Inf,-1]" = "#ce472e",
    "(-1,-0.75]" = "#f05336",
    "(-0.75,-0.5]" = "#faa224",
    "(-0.5,-0.25]" = "#ffd73e",
    "(-0.25,0]" = "#efe3be",
    "(0,0.25]" = "#c6e3bb",
    "(0.25,0.5]" = "#a3d393",
    "(0.5,0.75]" = "#64bc52",
    "(0.75,Inf]" = "#0099dc"
  )) +
  annotate("segment", x= 1.5, xend = 1.5, y = 0.35, yend = 0.15, 
           color = "black", linewidth = 0.5) +
  annotate("segment", x = 5.5, xend = 5.5, y = 0.35, yend = 0.15, 
           color = "black", linewidth = 0.5) +
  annotate("segment", x = 8.5, xend = 8.5, y = 0.35, yend = 0.15, 
           color = "black", linewidth = 0.5) +  
  
  annotate("text", x = 1.5, y = -0.1, label = "-1%", size = 2.4) +
  annotate("text", x = 5.5, y = -0.1, label = "0%", size = 2.4) +
  annotate("text", x = 8.5, y = -0.1, label = "0.75%", size = 2.4) +
  
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_fixed(ratio = 1, clip = "off") +
  theme_void() +
  theme(
    legend.position = "none",
    plot.margin = margin(5, 5, 20, 5)
  )

p2
```

## Join plots

```{r}
ggdraw() +
  draw_plot(p1) +
  draw_plot(p2, x = 0.65, y = 0.75, width = 0.25, height = 0.20)
```


## Improvement
new plot data
```{r}
bar_plot_new <- percentage_payroles |> 
  pivot_longer(
    cols = -date,
    names_to = "sector",
    values_to = "percentage_change"
  ) |> 
  filter(date >= "2008-02-01" & date <= "2017-03-01") 

```



maybe change percentage_payroles to include zero_data
```{r}
ggplot(bar_plot_new, aes(x = date, y = percentage_change, fill = sector)) +
  geom_col(position = "stack", width = 30) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 9, angle = 0, hjust = 0.5),
    axis.title.y = element_text(size = 9),
    
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    
    plot.margin = margin(10, 10, 10, 10),
    legend.position = "right",
    legend.text = element_text(size = 7),
    legend.key.size = unit(0.4, "cm")
  ) +
  labs(x = NULL,
       y = "Total Percentage Change",
       fill = "Sector",
       title = "Employment Changes by Sector Over Time"
       )
  
```


Small plots
Create a named vector with shaded colors for each sector
```{r}
sector_colors <- c(
  # Goods-producing (RED shades)
  "mining_and_logging" = "#245501",             
  "construction" = "#538d22",                     
  "durable_goods_manufacturing" = "#73a942",      
  "nondurable_goods_manufacturing" = "#aad576",   
  
  #
  "transportation_and_warehousing" = "#ffe5ec",   
  "wholesale_trade" = "#ffb3c6",             
  "utilities" = "#e0aaff",                     
  "information" = "#c77dff",                   
  "financial_activities" = "#9d4edd",             
  "professional_and_business_services" = "#5a189a", 
  "retail_trade" = "#240046",  
  "education" = "#03045e",                        
  "health_care_and_social_assistance" = "#023e8a", 
  "accommodation" = "#0096c7",      
  "restaurants_and_bars" = "#48cae4",     
  "other_services" = "#90e0ef", 
  "arts_entertainment_and_rec" = "#caf0f8", 
  #-----------------------------------------------------------
  # Trade & Transportation (BLUES - 4 sectors)
  "wholesale_trade" = "#08519C",
  "retail_trade" = "#3182BD",
  "transportation_and_warehousing" = "#48cae4",
  "utilities" = "#9ECAE1",
  
  # Business & Financial Services (PURPLES - 3 sectors)
  "information" = "#5a189a",
  "financial_activities" = "#9d4edd",
  "professional_and_business_services" = "#e0aaff",
  
  # Consumer & Social Services (TEALS/GREENS - 6 sectors)
  "education" = "#ffea00",
  "health_care_and_social_assistance" = "#ffb700",
  "accommodation" = " n#ffaa00",
  "restaurants_and_bars" = "#A1D99B",
  "arts_entertainment_and_rec" = "#C7E9C0",
  "other_services" = "#E5F5E0"
  
  # Government (GREEN shades)
  "federal_government" =  "#8B0000",               
  "state_government" = "#ad2e24",                 
  "local_government" = "#c75146"                  
)

```



Goods producing industries
```{r}
goods_plot <- bar_plot_new|> 
  filter(sector %in% c("mining_and_logging", "construction", 
                        "durable_goods_manufacturing", "nondurable_goods_manufacturing")
         )|> 
  ggplot(aes(x = date, y = percentage_change, fill = sector)) +
  geom_col(position = "stack", width = 25) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2009:2017, "-01-01"))), 
           color = "grey60", linewidth = 0.3, linetype = "dotted") +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0, 0), position = "top") +
  scale_y_continuous(expand = c(0, 0),
                      limits = c(-8, 15)) +
  scale_fill_manual(values = sector_colors,
                    labels= c("construction" = "Construction",
             "durable_goods_manufacturing" = "Durable Goods Manufact.",
             "mining_and_logging" = "Mining & Logging",
             "nondurable_goods_manufacturing" = "Nondurable Goods Manufact.")) +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
  ) +
  labs(title = "Goods-producing industries", x = NULL, y = NULL) +
  theme(legend.position = "bottom", legend.text = element_text(size = 8))

goods_plot
```


Private service provinding industries
```{r}

library(RColorBrewer)

service_plot <- bar_plot_new|> 
  filter(sector %in% c("transportation_and_warehousing", "wholesale_trade", "utilities", "information",
                       "financial_activities", "professional_and_business_services", "retail_trade", 
                       "education", "health_care_and_social_assistance", "accommodation", 
                       "restaurants_and_bars", "other_services","arts_entertainment_and_rec")
         ) |> 
  ggplot(aes(x = date, y = percentage_change, fill = sector)) +
  geom_col(position = "stack", width = 25) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2009:2017, "-01-01"))), 
           color = "grey60", linewidth = 0.3, linetype = "dotted") +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0, 0), position = "top") +
  scale_y_continuous(expand = c(0, 0),
                      limits = c(-8, 15)) +
  scale_fill_manual(values = sector_colors,
                    labels= c("transportation_and_warehousing" = "Transportation & warehousing",
                              "wholesale_trade" = "Wholesale trade",
                              "utilities" = "Utilities",
                              "information" = "Information",
                              "financial_activities" = "Financial activities",
                              "professional_and_business_services" = "Professional & business serv.",
                              "retail_trade" = "Retail trade",
                              "education" = "Education",
                              "health_care_and_social_assistance" = "Healthcare & social assistance",
                              "accommodation" = "Accommodation",
                              "restaurants_and_bars" = "Restaurants & bars",
                              "other_services" = "Other services",
                              "arts_entertainment_and_rec" = "Arts, entertainment & rec.")) +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
  ) +
  labs(title = "Goods-producing industries", x = NULL, y = NULL) +
  theme(legend.position = "right", legend.text = element_text(size = 8))

service_plot


```

Government
```{r}
gov_plot <- bar_plot_new|> 
  filter(sector %in% c("federal_government", "state_government", "local_government")
         )|> 
  ggplot(aes(x = date, y = percentage_change, fill = sector)) +
  geom_col(position = "stack", width = 25) +
  geom_vline(xintercept = as.numeric(as.Date(paste0(2009:2017, "-01-01"))), 
           color = "grey60", linewidth = 0.3, linetype = "dotted") +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", expand = c(0, 0), position = "top") +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(-8, 15)) +
  scale_fill_manual(values = sector_colors,
                    labels= c("federal_government" = "Federal gov.", 
                              "state_government" = "State gov.", 
                              "local_government" = "Local gov.")
                    ) +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
  ) +
  labs(title = "Goods-producing industries", x = NULL, y = NULL) +
  theme(legend.position = "bottom", legend.text = element_text(size = 8))

gov_plot
```


Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


