---
title: "Course contents"
---

## Chapter 0. Course Description

- Big picture
- Background and goals
- Course contents and schedule
- Logistics
- Communication and tutoring
- Course assessment
- [Example of final project](slides/ch0.html#8) [[article](projects/gapminder/)]
- [Bibliography and other resources](resources.html)

<aside class="slides">
```{r, echo=FALSE}
xaringanExtra::embed_xaringan("slides/ch0.html", "4:3")
```
</aside>

## Chapter 1. Fundamentals of Graphical Practice

- [Introduction to DataViz: The What, the Why and the How](slides/ch1.html#3)
  - What is a graph?
  - Acclaimed historic examples
  - Why graphs?
  - How to make graphs?
  - Principles of graphical excellence
- [Graphical Integrity: How to Lie with Graphs](slides/ch1.html#44)
  - Bad charts
  - Sources of bad charts
  - 12 rules to display data badly
  - Chart rethinking (some exercises)
- [Graphical Perception: Theory, Experimentation and its Application to Data Display](slides/ch1.html#75)
  - Introduction
  - Elements of visual encoding
  - Expressiveness principle
  - Effectiveness principle
  - Channel effectiveness
- [Principles of Graphical Representation: Some Useful General Guidelines](slides/ch1.html#112)
  - Attitude
  - General Guidelines

<aside class="slides">
```{r, echo=FALSE}
xaringanExtra::embed_xaringan("slides/ch1.html", "4:3")
```
</aside>

## Chapter 2. The Grammar of Graphs in R

- [Building Graphs Layer by Layer](slides/ch2.html#3)
  - An object-oriented graphics system
  - About `ggplot2`
  - `ggplot2` basics
  - Aesthetic specifications
  - Individual geoms
  - Collective geoms
  - Geom vs. Stat
  - [Tutorial 01](tutorials/01/)
- [Scales and Guides](slides/ch2.html#40)
  - Scale specification
  - [Tutorial 02](tutorials/02/)
- [Coordinate Systems](slides/ch2.html#45)
  - Cartesian coordinates
  - Other coordinates
  - [Tutorial 03](tutorials/03/)
- [Faceting](slides/ch2.html#50)
  - Facet specification
  - [Tutorial 04](tutorials/04/)
- [Themes](slides/ch2.html#54)
  - Theme specification
  - [Tutorial 05](tutorials/05/)

<aside class="slides">
```{r, echo=FALSE}
xaringanExtra::embed_xaringan("slides/ch2.html", "4:3")
```
</aside>

## Chapter 3. Data Visualization in R

- [Directory of Visualizations](slides/ch3.html#3)
  - Distribution
  - Correlation
  - Ranking
  - Part of a Whole
  - Evolution
  - Map
  - Flow
  - Other Resources

<aside class="slides">
```{r, echo=FALSE}
xaringanExtra::embed_xaringan("slides/ch3.html", "4:3")
```
</aside>
