---
title: "Course contents"
---

## [Chapter 0](slides/ch0.html). Course Description

- Big picture
- Background and goals
- Course contents and schedule
- Logistics
- Communication and tutoring
- Course assessment
- [Example of final project](slides/ch0.html#8) [[article](projects/gapminder/)]

<aside class="slides">
```{r, echo=FALSE}
xaringanExtra::embed_xaringan("slides/ch0.html", "4:3")
```
</aside>

## [Chapter 1](slides/ch1_1.html). Fundamentals of Graphical Practice

- [Introduction to DataViz](slides/ch1_1.html#3)
- [Graphical Integrity](slides/ch1_2.html#3)
- [Graphical Perception](slides/ch1_3.html#3)
- [Principles of Graphical Representation](slides/ch1_4.html#3)

<aside class="slides">
```{r, echo=FALSE}
xaringanExtra::embed_xaringan("slides/ch1_1.html", "4:3")
```
</aside>

## [Chapter 2](slides/ch2.html). The Grammar of Graphs in R

- [Building Graphs Layer by Layer](slides/ch2.html#3)
- [Scales and Guides](slides/ch2.html#40)
- [Coordinate Systems](slides/ch2.html#45)
- [Faceting](slides/ch2.html#50)
- [Themes](slides/ch2.html#54)
- [Annotations](slides/ch2.html#57)
- [Arranging Plots](slides/ch2.html#66)

<aside class="slides">
```{r, echo=FALSE}
xaringanExtra::embed_xaringan("slides/ch2.html", "4:3")
```
</aside>

## [Chapter 3](slides/ch3_1.html). Data Visualization in R

- [Distribution](slides/ch3_1.html#3) Graphs
- [Correlation](slides/ch3_2.html#3) Graphs
- [Ranking](slides/ch3_3.html#3) Graphs
- [Part of a Whole](slides/ch3_4.html#3) Graphs
- [Evolution](slides/ch3_5.html#3) Graphs
- [Maps](slides/ch3_6.html#3)
- [Flow](slides/ch3_7.html#3) Charts
- [Other Resources](slides/ch3_8.html#3)

<aside class="slides">
```{r, echo=FALSE}
xaringanExtra::embed_xaringan("slides/ch3_1.html", "4:3")
```
</aside>

## [Chapter 4](slides/ch4_1.html). Data Communication in R

- [Version control](slides/ch4_1.html#3)
- [R Markdown](slides/ch4_2.html#3)
  - Documents and notebooks
  - Presentations
  - Other formats
  - Dashboards
  - Shiny

<aside class="slides">
```{r, echo=FALSE}
xaringanExtra::embed_xaringan("slides/ch4_1.html", "4:3")
```
</aside>
