---
title: "Course contents"
---

# Chapter 0. Course Description

- Big picture
- Background and goals
- Course contents and schedule
- Logistics
- Communication and tutoring
- Course assessment
- [Example of final project](slides/ch0.html#8) [[article](projects/gapminder/)]
- [Bibliography and other resources](slides/ch0.html#26)

<aside class="slides">
```{r, echo=FALSE}
xaringanExtra::embed_xaringan("slides/ch0.html", "4:3")
```
</aside>

# Chapter 1. Fundamentals of Graphical Practice

- [Introduction to DataViz: The What, the Why and the How](slides/ch1.html#3)
  - What is a graph?
  - Acclaimed historic examples
  - Why graphs?
  - How to make graphs?
  - Principles of graphical excellence
- [Graphical Integrity: How to Lie with Graphs](slides/ch1.html#44)
  - Bad charts
  - Sources of bad charts
  - 12 rules to display data badly
  - Chart rethinking (some exercises)

<aside class="slides">
```{r, echo=FALSE}
xaringanExtra::embed_xaringan("slides/ch1.html", "4:3")
```
</aside>
