---
title: "The world is progressively embracing democracy"
description:|
  A reproduction of the Our World in Data visualization, which is based on the classification and assessment of countries' 
  political regimes by Skaaning et al. (2015).
  categories:"2023"
author: Sofia Villamil Erosa
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    self_contained: false
    toc: true
---

## Democracy

The graphic I choose shows how democracy has spread across countries, how it differs between them, and whether we are moving 
towards a more democratic world. The variables denoted in the chart shows the number of electoral democracies in the world based 
on the Lexical Index of Electoral Democracy-classification from Skaaning et al. (2015).

What is the Lexical Index?

The Lexical Index definition attained from Our World in Data establishes that it is a classification system that categorizes 
political systems or regimes based on easily observable characteristics. It aims to avoid difficult evaluations by experts or 
researchers and relies on its own teams to assess specific criteria. These criterions are used to identify democracies and 
non-democracies. The index is able to track political systems over time, measured since the year 1789, and includes 242 countries.
 The index prioritizes straightforward and objective factors, making it a tool for quick assessments and comparisons.

### Definitions

**Types of Democracy in the Lexical Index:**

1.  **Non-Electoral Autocracy:** A political system characterized by a single entity or a small group holding power without 
elections, often with a single autocrat or ruling elite.

2.  **One-Party Autocracy:** A system where a single political party maintains absolute control over the government, suppressing 
the opposition and competitive elections.

3.  **Multi-Party Autocracy without Elected Executive:** A political system with multiple parties, but the executive leader is 
not elected democratically.

4.  **Multi-Party Autocracy:** Features multiple political parties, but the political system still retains autocratic elements.

5.  **Exclusive Democracy:** Inclusive of elections but excludes certain groups or individuals from political participation.

6.  **Male Democracy:** A democracy where only male citizens are allowed to participate in elections.

7.  **Electoral Democracy:** A political system with regular elections where citizens can choose their leaders, but this may not 
guarantee full democratic rights.

8.  **Polyarchy:** A comprehensive democracy with competitive elections, civil liberties, political pluralism, and broad 
participation.

These definitions reflect the categorization of political systems within the Lexical Index, based on specific characteristics and 
levels of inclusiveness attained from Skaaning et al. (2015), with "Polyarchy" representing the highest level of democracy and 
inclusivity.

### Getting the data

The dataset, obtained from Our World in Data and based on research by Skaaning et al. (2015), categorizes the world's nations 
into democracies and autocracies. This graphic dynamically captures the shifts in these counts from 1789 through 2022, offering 
insights into the evolving landscape of global political systems nowadays. All the data needed for this chart can be downloaded 
directly from the Our World in Data website in the topic Democracy 
(<https://ourworldindata.org/grapher/countries-democracies-autocracies-lexical>).

#### Library and the data set with the frequency variables

```{r}
data <- read.csv("Data World csv.csv", 
                 sep = ";", 
                 dec = ",")
library(ggplot2)
library(dplyr)
library(tidyr)
```

### Building the chart

#### Data

The data scoured from Our World in Data needed to be converted to frequency tables . The graph's vertical axis, which I aimed to 
replicate, utilized cumulative frequency as a metric, necessitating the data's conversion. Subsequently, I identified and chose 
the specific columns needed for the replication, resulting in the creation of a new data-set named *df_reduced*.

```{r}
df_reduced <- data %>%
  select(Year, Frec.Non.electoral.autocracies, Frec.One.party.autocracies, Frec.Multi.party.autocracies.without.elected.executive,
         Frec.Multi.party.autocracy, Frec.Exclusive.democracies, Frec.Male.democracies, Frec.Electoral.democracies, 
         Frec.Polyarchies) 

#FREC - refers to the frecuency of the


custom_order <- c("Frec.Non.electoral.autocracies", "Frec.One.party.autocracies", 
"Frec.Multi.party.autocracies.without.elected.executive",
                  "Frec.Multi.party.autocracy", "Frec.Exclusive.democracies", "Frec.Male.democracies", 
                  "Frec.Electoral.democracies", "Frec.Polyarchies")
```

To enhance accessibility and facilitate the chart-building process, I converted the new dataset into tidy data. During this 
conversion, several changes were implemented, and intervals were established for the axis.

```{r}
df_long <- df_reduced %>%
  pivot_longer(cols = -Year, names_to = "Democracy_Type", values_to = "n")

df_long <- df_long %>%
  mutate(Democracy_Type = factor(Democracy_Type, levels = custom_order))

intervalos_x <- c(1789, 1850, 1900, 1950, 2000, 2022)

intervalos_y <- seq(0, 100, by = 0.2)  

```

Due to each variable having an associated color. The following color vector was created for each value.

```{r}
colores_democracias <- c( "Frec.Non.electoral.autocracies" = "#d73027", "Frec.One.party.autocracies" = "#f46d43", 
"Frec.Multi.party.autocracies.without.elected.executive" = "#fdae61", "Frec.Multi.party.autocracy" = "#fee090", 
"Frec.Exclusive.democracies" = "#e0f3f8", "Frec.Male.democracies" = "#abd9e9", "Frec.Electoral.democracies" = "#74add1", 
"Frec.Polyarchies" = "#4575b4" )
```

#### Starting with the graphic

To construct the visual representation, I utilized the *geom_area* function from *ggplot*, tailored for generating 
**Stacked Area Charts.**

```{r}
plot <- ggplot(df_long, aes(x = Year, y = n, fill = Democracy_Type)) +
  geom_area(alpha = 0.8, size = 0.2, aes(color = Democracy_Type)) +
  labs(x = NULL, y = NULL, fill = NULL) +
  scale_x_continuous(limits = c(1789, 2085), breaks = intervalos_x, minor_breaks = NULL) +
  scale_y_continuous(labels = scales::percent_format(scale = 100), breaks = intervalos_y) +
  scale_fill_manual(values = colores_democracias) +
  scale_color_manual(values = colores_democracias) +
  coord_cartesian(expand=0) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(color="#5b5b5b", size = 10, family = "Lato", angle = 0, hjust = c(0.1, 0.3, 0.5, 0.5, 0.5, 1),
     margin = margin(t = 0)),
    axis.ticks.x = element_line(color="#5b5b5b", size = 0.2),
    axis.ticks.length = unit(0.1, "cm"),
    legend.position = "none",
    panel.grid = element_blank(),
    axis.text.y = element_text(color="#5b5b5b", size = 10, family = "Lato", margin = margin(r = 0))
  ) +
  ggtitle("Countries that are democracies and autocracies, World", subtitle = "Political regimes based on the classification and 
  assessment by Skaaning et al. (2015).") +
  theme(plot.title = element_text(color="#5b5b5b", family = "serif", size = 15, face = "bold", hjust = -0.27),
        plot.subtitle = element_text(color="#5b5b5b", family = "Lato", size = 9, hjust = -0.24))
plot
```

I employed the intervals created during the data organization for both the x-axis and the y-axis in constructing the graphic. 
Given that the y-axis represents frequencies in percentage, the code was modified to reflect this in the visualization. The year 
limits were adjusted to fit the dimensions required for displaying custom labels for each democratic type. While the code's limits
 span from 1789 to 2085, the graphic displays the period from 1789 to 2022 on the x-axis.

To achieve a custom separation between each year on the x-axis, adjustments were made in the theme section, specifically in 
*axis.text.x*, where the spacing between each year was manually configured in order to be exactly the distance displayed 
in the graphic of reference.

#### Final touches

I generated custom labels with corresponding colors and segments, mirroring the structure observed in the graphic that I aimed 
to replicate.

```{r}
leyenda_custom <- data.frame(
  Democracy_Type = levels(df_long$Democracy_Type),
  x = rep(2022, length(levels(df_long$Democracy_Type))),
  y = seq(0, 1, length.out = length(levels(df_long$Democracy_Type))),
  label = c("Polyarchies",
  "Electoral
democracies",
  "Male 
democracies",
  "Exclusive 
democracies",
    "Multi-party
autocracy",
    "Multi-party
autocracies
without elected
executive",
    "One-party
autocracies",
    "Non-electoral
autocracies"
  ),
  distancia_etiqueta = c(11, 11, 11, 21, 4, 31, 4, 8),
  distancia_segmento = c(10, 10,  10,  20,  3,  30,  3,  7),
  angulo_segmento = c(0, 0, 0, 0, 0, 0, 0, 0),  
  desplazamiento_vertical_segmento = c(0.20, 0.33, 0.338, -0.16, 0.15, 0.09, 0.01, -0.04),
desplazamiento_vertical_etiqueta = c(0.20, 0.33, 0.338, -0.16, 0.15, 0.09, 0.01, -0.04),
extra =c(0, 0, 0, 0, 0, 0, 0, 0)
)
Democracy_Type_corregido <- rev(leyenda_custom$Democracy_Type)

```

To replicate the specific placement of labels and segments in the target graphic, I modified the code to permit manual 
adjustments for the angle, vertical, and horizontal placement of both text and segments. Thus, *distancia_segmento* makes 
reference to the horizontal placement of the segment, *angulo_segmento* is for the angle and *desplazamiento_vertical_segmento* 
is for the vertical placement.

Furthermore, I introduced a custom legend to the plot and fine-tuned the size parameters to precisely match the dimensions of 
the desired graphic.

```{r}
plot2 <- plot +
  geom_text(
    data = leyenda_custom,
    aes(x = x + distancia_etiqueta, y = y + desplazamiento_vertical_etiqueta, label = label, color = Democracy_Type_corregido),
    size = 2.5,
    family = "Lato",
    hjust = 0
  ) +
  geom_segment(
    data = leyenda_custom,
    aes(x = x + distancia_segmento , xend = x+1+extra, y = y + desplazamiento_vertical_segmento, yend = y + angulo_segmento + 
    desplazamiento_vertical_segmento, color = "#5b5b5b"),
    size = 0.5
  ) +
  theme(
    plot.margin = margin(0.3, 0.3, 0.3, 0.3, "cm")
  )

plot2


```

### Improving the chart

Upon careful consideration, I determined that there was no apparent need for enhancements in the replicated graphic. Consequently,
 I opted to generate a new visual representation utilizing the same dataset.

### Building the new chart

In an effort to retain as much of the original code as possible, I generated a new custom color palette to address perceived 
deficiencies in the previous one. This adjustment was made to enhance visibility, ensuring that all colors are easily discernible.
 Additionally, I maintained the code for creating custom labels in the process as I wanted to highlight that code and expand the
  use of it to improve the new graphic.

```{r}
colores_democracias <- c(
  "Frec.Non.electoral.autocracies" = "#081d58",
  "Frec.One.party.autocracies" = "#253494",
  "Frec.Multi.party.autocracies.without.elected.executive" = "#225ea8",
  "Frec.Multi.party.autocracy" = "#41b6c4",
  "Frec.Exclusive.democracies" = "#7fcdbb",
  "Frec.Male.democracies" = "#c7e9b4",
  "Frec.Electoral.democracies" = "#1d91c0",
  "Frec.Polyarchies" = "black"
)
leyenda_custom <- data.frame(
  Democracy_Type = levels(df_long$Democracy_Type),
  x = rep(2022, length(levels(df_long$Democracy_Type))),
  y = seq(0, 1, length.out = length(levels(df_long$Democracy_Type))),
  label = c(
    "Polyarchies",
    "Electoral democracies",
    "Male democracies","Exclusive democracies",
    "Multi-party autocracy",
    "Multi-party autocracies without elected executive",
    "One-party autocracies",
    "Non-electoral autocracies"
  ),
  distancia_etiqueta = c(3, -20, 3, 3, -60, -220  , 3, -220),
  distancia_segmento = c(1, 1, 1, 1, 1, 1, 1, 1),
  angulo_segmento = c(0, 0, 0, 0, 0, 0, 0, 0),
  desplazamiento_vertical_segmento = c(0, 0, 0, 0, 0, 0, 0, 0),
  desplazamiento_vertical_etiqueta = c(0.38, 0.17,-0.28, -0.27, -0.59, -0.32, -0.76, 0),
  extra = c(0, 0, 0, 0, 0, 0, 0, 0)
)

Democracy_Type_corregido <- rev(leyenda_custom$Democracy_Type)
```

The chosen type of chart for improvements was a **Parallel Plot** utilizing the *geom_line* function from *ggplot*. I chose this
 approach as it provides a clearer representation of the evolution of each democracy type throughout the analysis period.

In trying to create the new graphic, I aimed for a fresh look while retaining the essence of the original one I sought to 
replicate. To achieve this, I employed much of the existing code I had created for the previous graphic. Additionally, I 
introduced some modifications to add distinctive elements that I deemed necessary, resulting in a new and improved interpretation 
of the same data set.

```{r}
plot_parallel_custom <- ggplot(df_long, aes(x = Year, y = n, color = Democracy_Type)) +
  geom_line(aes(group = Democracy_Type), alpha = 0.8, size = 1) +
  scale_x_continuous(limits = c(1789, 2065), breaks = intervalos_x, minor_breaks = NULL) +
  scale_y_continuous(labels = scales::percent_format(scale = 100), breaks = intervalos_y) +
  geom_text(
    data = leyenda_custom,
    aes(x = x + distancia_etiqueta, y = y + desplazamiento_vertical_etiqueta, label = label, color = Democracy_Type_corregido),
    size = 2.8,
    family = "Playfair Display ExtraBold",
    hjust = 0) +
  scale_color_manual(values = colores_democracias) +
  labs(
    x = "Years",
    y = "Percentage of Countries",
    title = "Countries that are democracies and autocracies, World.",
    subtitle = "Political regimes based on the classification and assessment by Skaaning et al. (2015)."
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(color = "#252525", family = "serif", size = 15, face = "bold", hjust = -0.1, vjust = 0),
    axis.title.x = element_text(color = "#252525", family = "serif", size = 12, hjust = 0.44, vjust = -1),   
    axis.title.y = element_text(color = "#252525", family = "serif", size = 12,vjust = 3), 
    plot.subtitle = element_text(color = "#252525", family = "Lato", size = 9, hjust = -0.1, vjust = 0, face = "italic"),
    axis.text.x = element_text( color = "#5b5b5b", size = 9, family = "Playfair Display ExtraBold", vjust = 0 , 
    hjust = c(0, 0, 0, 0, 0.5, 1), margin = margin(t = -9)),
  axis.text.y = element_text(color = "#5b5b5b", size = 9, family = "Playfair Display ExtraBold", vjust = 0, hjust = 0,
   margin = margin(r = -22 )
    )) +
  theme(
    plot.margin = margin(0.4, 0.4, 0.5, 0.7, "cm"), 
  ) +
  guides(color = "none")

plot_parallel_custom
```

Recognizing the need for enhanced interpretability, I decided to include legends for both the x-axis and y-axis, making it easier 
to understand the various measures depicted in the plot. To improve readability, I made adjustments such as changing the font 
color to black. Moreover, I relocated the labels inside the graphic instead of just using segments, believing that this 
modification would enhance comprehension of each type of democracy and its evolution over time.
