---
title: "02. Scales and Guides"
description: |
  Scales are responsible for turning your mapping from data to aesthetics into
  something that you can see, and they also provide the guides that let you
  interpret the plot: axes and legends. Scales are automatically inferred based
  on the input data, and guides are added as well based on the layers in your
  plot, which is convenient and saves a lot of time. However, sometimes it is
  required to fine-tune how the mapping is performed, or how the guide should
  look like. This tutorial explores the tools available for this.
categories: "Chapter 2"
author:
  - name: Iñaki Ucar
    affiliation: Department of Statistics
    orcid_id: 0000-0001-6403-5550
date: 2022-10-05
output:
  distill::distill_article:
    self_contained: false
    toc: true
preview: scale-guides.png
draft: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
registerS3method("knit_print", "data.frame", function(x, ...) {
  rmarkdown:::knit_print.data.frame(rmarkdown::paged_table(x))
}, envir = asNamespace("knitr"))
```

Source: [`_tutorials/02/02.Rmd`](https://github.com/IBiDat/dataviz/blob/main/_tutorials/02/02.Rmd)

## Scale specification

While **geom**s represent visual objects (points, lines...), a **scale** is a procedure that actually *performs* the mapping of data attributes into the properties of those visual objects (position, color, size...). On the other hand, a **guide** is a visual cue that allows *us* to perform the inverse procedure: converting their properties back to data.

```{r}
library(ggplot2)

# convert number of cylinders to factor (only 3 levels)
mtcars$Cylinders <- as.factor(mtcars$cyl)

ggplot(mtcars) +
  aes(mpg, wt) +
  geom_point(aes(color=Cylinders))
```

![Common components of axes and legends. Figure from the [ggplot2 book](https://ggplot2-book.org/scales-guides.html).](scale-guides.png){.external}

In the example above, two position scales (**x** and **y**)

-   detected that the data (`mpg` and `wt`) is **continuous**;
-   set proper limits to the axes according to the range of data values;
-   assigned a position in the chart space to each point according to such Cartesian coordinates;
-   added the corresponding guides to the axes, including
    -   nicely and evenly spaced breaks, with tick marks and labels,
    -   centered axis labels with the name of the variables.

Additionally, a **color** scale

-   detected that the data (`Cylinders`) is **discrete** (a categorical variable, a factor);
-   assigned a color to each category following a default hue-based scale;
-   added the corresponding guide to the legend, including
    -   set of keys (colors),
    -   set of labels (categories in the data),
    -   legend title with the name of the variable.

In other words, the plot above is equivalent to the following:

```{r}
ggplot(mtcars) +
  aes(mpg, wt) +
  geom_point(aes(color=Cylinders)) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_color_discrete()
```

The last three instructions are just inferred from the data and automatically added. We will only need to specify the scales manually whenever we want to override any of the defaults provided.

There are three fundamental scale types:

-   **continuous**, for numerical quantities;
-   **discrete**, for categorical variables;
-   **binned**, which is a special type for numerical quantities that first discretize the data.

Based on these, the fundamental scales, in general, are constructed as follows:

    scale_<aes>_<type>()

For example:

-   For aesthetic `x` and a continuous variable: `scale_x_continuous()`.
-   For aesthetic `color` and a categorical variable: `scale_color_discrete()`.

Then there are a bunch of special scales that are shortcuts for specific common tasks, such as `scale_x_log10()`, which is a continuous scale that also applies a logarithmic transformation, or `scale_color_brewer()`, which provides a bunch of discrete color schemes from [ColorBrewer](https://colorbrewer2.org/).

#### Checkpoint 1

1.  Take the first example and assign the original `cyl` variable to color instead of `Cylinders`. What happens? Why? What is the scale function that is being applied here? What happens if we try to force `scale_color_discrete()` in this setting? Why?

2.  Read the documentation for [`scale_x_continuous()`](https://ggplot2.tidyverse.org/reference/scale_continuous.html). How can you change the name of the axis? What happens if we manually specify the same scale twice (try adding different names)? Why?

## Scale names

As you may have discovered in the previous section, all scales have a `name` argument that is used for the axis label (for position scales) or legend title (for other scales), and which by default prints the name of the variable. There is however a shortcut, a more direct way of setting these labels, via the `labs()` function.

```{r}
ggplot(mtcars) +
  aes(mpg, wt) +
  geom_point(aes(color=Cylinders)) +
  labs(
    x = "Fuel consumption in miles per gallon",
    y = "Weight in thousands of pounds",
    color = "Number of cylinders"
  )
```

In these strings, you can use `\\n` to insert a line break, but you can also supply mathematical expressions and other formatting options wrapped in `quote()`. See `?plotmath` for documentation on which expressions can be used and how they are interpreted.

```{r}
x <- seq(from = -2, to = 2, by = .01)
ggplot(data.frame(x=x, y=x^3)) +
  aes(x, y) +
  geom_path() +
  labs(y = quote(f(x) == x^3))
```

#### Checkpoint 2

1.  How can you remove an axis label, or a legend title?

2.  Read the documentation for [`labs()`](https://ggplot2.tidyverse.org/reference/labs.html). Investigate what other labels, apart from the aesthetics, can be set with this function. Experiment with them.

3.  Read the documentation for `?plotmath`. How can you set a title with some words in bold, others in italics...? How can you wrap a very long subtitle with line breaks?

## Limits

## Breaks and labels

## Transformations

## Guides
