---
title: "03. Coordinate Systems"
description: |
  The meaning of position aesthetics and how they produce a 2D position on the
  plot depend on the coordinate system. This is also responsible for drawing
  the axes and panel backgrounds (grid lines, etc.). As with scales, Cartesian
  coordinates are applied by default unless stated otherwise, which preserves
  the common meaning of x and y.
categories: "Chapter 2"
author:
  - name: Iñaki Ucar
    affiliation: Department of Statistics
    orcid_id: 0000-0001-6403-5550
date: 2022-10-05
output:
  distill::distill_article:
    self_contained: false
    toc: true
preview: preview.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE)
registerS3method("knit_print", "data.frame", function(x, ...) {
  rmarkdown:::knit_print.data.frame(rmarkdown::paged_table(x))
}, envir = asNamespace("knitr"))
```

Source: [`_tutorials/03/03.Rmd`](https://github.com/IBiDat/dataviz/blob/main/_tutorials/03/03.Rmd)

## Cartesian coordinates

Most of the plots we know and love use Cartesian coordinates, where the position of an element is given by orthogonal distances, `x` and `y`, to an origin.

```{r}
library(ggplot2)

ggplot(mpg) +
  aes(displ, hwy) +
  geom_point(aes(color=class))
```

Even when one of the axis is a categorical variable, the same concept of "distance to an origin" applies, because internally those categories are mapped to positions 1, 2, 3...

```{r}
ggplot(mpg) +
  aes(displ, class) +
  geom_col() +
  geom_text(aes(label=class), data.frame(
    displ = 200, class = c(1, 2, 3, 4, 5, 6, 7)
  ))
```

As with scales, `coord_cartesian()` is the default, and it does not need to be explicitly specified to produce a plot. However, this function has some useful options that makes some things easier. Particularly, you may recall from the previous tutorial that zooming in by setting the limits of a scale by default censors points that are out of bounds (meaning that they are marked as missing with a `NA`). For scales, an additional argument `oob` must be specified in order to keep those points (e.g. for smoothing calculations). In contrast, `coord_cartesian()` provides limits that do not censor data. Therefore, the following plots are equivalent:

```{r}
p <- ggplot(mpg) +
  aes(displ, hwy) +
  geom_point(aes(color=class)) +
  geom_smooth()

p +
  scale_x_continuous(limits=c(3, 4), oob=scales::oob_keep) +
  scale_y_continuous(limits=c(20, 30), oob=scales::oob_keep)

p +
  coord_cartesian(xlim=c(3, 4), ylim=c(20, 30))
```

#### Checkpoint 1

1.  Function `coord_flip()` comes in handy to exchange the `x` and `y` axes. Take the first example and add `geom_smooth(method="lm")`. What is the difference between adding `coord_flip()` and directly exchanging the `x` and `y` mapping in the `aes()`? Why?

2.  Take the first example and add `coord_fixed()`. Then use variable `cty` instead of `displ` to compare. What does this function do? What could be a good use case for this?

## How to make a pie

A pie chart is just a single stacked bar in polar coordinates. For instance, take this one:

```{r}
p <- ggplot(mpg) +
  aes("cars", fill=class) +
  geom_bar()
p
```

Then, we just need to fold it using `y` (the count produced by `geom_bar`) as the angle:

```{r}
p + coord_polar(theta="y")
```

So much effort for such a meaningless visualization. :)

## Transformations

Transformations can be performed at two levels: at the scale level, via the `trans` argument to the scale functions (or the dedicated scales such as `scale_x_log10()`), or at the coordinate system level, via `coord_trans()`. The only difference is that the first ones occur **before** any statistics are computed, while the second ones occur **after** those.

#### Checkpoint 2

1.  Take the first example and add `geom_smooth(method="lm")`. What is the difference between adding `scale_x_log10()` and adding `coord_trans(x="log10")`? Why?

## Maps

Finally, an important family of coordinate systems are map projections. Enter the fascinating world of how to map a sphere to a plane.

```{r}
p <- ggplot(map_data("world")) +
  aes(long, lat, group=group) +
  geom_path()

p # No projection specified, simply long, lat as x, y
p + coord_map(xlim=c(-180, 180)) # Mercator projection by default
p + coord_map("orthographic", xlim=c(-180, 180))
p + coord_map("sinusoidal", xlim=c(-180, 180))
```

See `?mapproj::mapproject` for more info about projections.

#### Checkpoint 3

1.  Click "Zoom" in the plot panel for each of the four plots proposed above. What happens if you stretch and shrink the width of the plot panel in each case? Why?
