---
title: "Seizures of Ammunition and Arms by Country 2017"
description: "A brief report based on UNs 2020 Global Study on Firearms Trafficking."
author: Mecki Hamed-Guerrero
categories: "2023"
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    self_contained: false
    toc: true
editor_options: 
  markdown: 
    wrap: 72
---

## Introduction

The illicit trade of weapons is a pervasive issue with far-reaching
consequences, intertwining elements of politics, society, and
international relations. The trafficking of arms fuels conflicts,
exacerbates human suffering, and poses a significant threat to global
security.

Arms trafficking knows no borders, infiltrating regions plagued by
political instability and social unrest. The consequences are dire, as
these weapons often fall into the hands of non-state actors,
perpetuating violence and hindering efforts for peace. The intertwined
nature of this problem reflects the complex web of international
relations, where the actions of one nation can have ripple effects
across the globe.

Tackling the issue of arms trafficking requires a multi-faceted
approach, involving collaboration between nations, strong regulations
and a commitment to addressing the root causes of conflict. It also
requires information to enable institutions to fully address the issue.
This is where data becomes necessary.

Establishing a relationship between the different types of arms
trafficking that exist and the conflicts that are currently taking place
can be very useful in understanding and addressing this phenomenon.
Creating a model that can explain the dynamics of arms (and components)
trafficking at the regional level would provide researchers and
decision-makers with a powerful tool to understand and prevent this
menace.

That is why the intention of this paper will be to plot the ratio of
ammunition seizures in relation to the number of arms seizures on a
world map. The motivation for this work lies in the intention to
identify regional trafficking dynamics around existing conflicts.

### The original graph

The original graph can be found on page 29 of the [UNODC Global Study on
Firearms Trafficking 2020](https://www.unodc.org/documents/data-and-analysis/Firearms/2020_REPORT_Global_Study_on_Firearms_Trafficking_2020_web.pdf).
The graph to be replicated is Figure 11.

The graph shows the relationship between ammunition seizures and weapons
seizures per country in 2017. The X-axis represents the countries in
which seizures occurred, and the Y-axis represents the ratio of seized
rounds of ammunition to seized arms (in a logarithmic scale). The graph
also has a line at the value 1 for the Y-axis, which indicates the
boundary between a positive and a negative ratio. Similarly, the graph
has a horizontal line above the median value (23), and a space dedicated
to the common range.

Countries above the ratio = 1 are those where ammunition trafficking is
higher than arms trafficking. This could indicate the presence of arms
from previous conflicts and belonging to already armed groups, which
require more ammunition than arms.

Conversely, where the ratio is \< 1, a process of internal rearmament
may be taking place. This is probably due to the involvement of third
states in the financing of paramilitary or terrorist groups.

![Original Source: [UNODC Global Study on Firearms Trafficking 2020](https://www.unodc.org/documents/data-and-analysis/Firearms/2020_REPORT_Global_Study_on_Firearms_Trafficking_2020_web.pdf).](original.png){.external}

## Working with the data

The database to be used comes from the UNODC (United Nations Office on
Drugs and Crime) and can be found on the following link: [UNODC Firearms
Trafficking Dataset](https://dataunodc.un.org/dp-firearms-arms-seized)

The dataset contains Categories relating to the 'Region', 'Subregion'
and 'Country'. Similarly, the dataset contains information relating to
the 'Source' that made the seizure and the year in which the seizure was
made (2010 to 2020). Also, the dataset contains a specific category,
'Indicator', which gives information about the type of seizure
(ammunition or arms). Finally, the dataset contains two more important
collumns: 'Dimension', which explains the reason of the seizured; and
'Category, which indicates the illegal use of the equipment seized.

However, the dataset lacks the data that the UNODC appears to use to
draft the aforementioned Global Study on Firearms Trafficking 2020. For
this reason, the scope of the work will be reduced to the information
available.

### Libraries and dataset

The following libraries will be utilized throughout the entire project,
encompassing both the replica and subsequent enhancements:

```{r}
library(datapasta)
library(tibble)
library(tidyr)
library(tidyverse)
library(dplyr)
library(readxl)
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(RColorBrewer)
```

Due to the fact that the first two columns of the chosen dataset
provides no information at all, we will load the .xlsx without them.

```{r echo=TRUE}

data <- read_excel("data_iafq_firearms_trafficking.xlsx", skip=2)
head(data)

```

### Data cleaning process

First of all:

The column "Year" represents the year of the record in the database.
Since our chart to be replicated relies on data for the year 2017, we
will select only records produced in that year.

```{r echo=TRUE}

data <- data[data$Year == 2017, ]
head(data)

```

Secondly:

The column "Indicator" indicates whether the data belongs to "Arms
seized" or "Parts and components seized". So we must filter the dataset
so that it contains only these two values:

```{r echo=TRUE}

data <- data |>
  filter(Indicator %in% c("Arms seized", "Ammunition seized"))
head(data)

```

Thirdly:

Since the columns "Dimension" and "Category" segment the total seizures
into different subdivisions, we are interested in showing only the sum
total of each of the columns. i.e. those values recorded as 'Total', in
both columns:

```{r echo=TRUE}

data <- data |>
  filter(Dimension == "Total" & Category == "Total")
head(data)

```

Fourth:

We convert the information in the "VALUE" column (total number of
seizures) to a numerical value:

```{r echo=TRUE}

data$VALUE <- as.numeric(data$VALUE)
head(data)

```

Fifthly:

We proceed to calculate the ratio of "Ammunition seized" to "Arms
seized", under the previously established conditions.

```{r echo=TRUE}

ratiocalc <- data |>
  group_by(data$Country) |>
  reframe(
    ratio = VALUE[Indicator == "Ammunition seized"] / VALUE[Indicator == "Arms seized"])
head(ratiocalc)

```

## Building the graph

### First replica

We proceed to create a first plot with the most basic information.

```{r echo=TRUE}

ggplot(ratiocalc, aes(x=reorder(`data$Country` , log10(ratio*1000)), y=log10(ratio*1000))) +
  geom_col() +
  scale_y_continuous(labels = function(x) 10^x/100) +
  labs(x = ' ',
       y= "Ratio of seized rounds of ammunition to seized arms
(logarithmic scale)"
  ) 

```

We then proceed to add both the value relative to the median and a
horizontal line at the value 1 for the Y-axis, which will allow us to
better differentiate countries above and below the ratio = 1.

Similarly, we can calculate the 'common range' and plot it on the graph.

```{r echo=TRUE}

mediana <- median(ratiocalc$ratio)

cuartil_ratio <- quantile(ratiocalc$ratio, c(0.25, 0.75))

rango_intercuartil_ratio <- cuartil_ratio[2] - cuartil_ratio[1]

ggplot(ratiocalc, aes(x=reorder(`data$Country` , log10(ratio*1000)), y=log10(ratio*1000))) +
  geom_col() +
  scale_y_continuous(labels = function(x) 10^x/100) +
  labs(x = ' ',
       y= "Ratio of seized rounds of ammunition to seized arms
(logarithmic scale)"
  ) +
  geom_hline(yintercept = log10(mediana*100), linetype = "dashed", color = "red") +
  geom_hline(yintercept = log10(100), linetype = "dashed", color = "red")  +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = log10(cuartil_ratio[1]*100), ymax = log10(cuartil_ratio[2]*100)),
            fill = "red", alpha = 0.003) 


```

From here we proceed to modify the aesthetic section of the graphic:

### Edition and color

The first modification consists of applying 'theme_minimal()' to the
graph. We will also add a title to it.

```{r echo=TRUE}

ggplot(ratiocalc, aes(x=reorder(`data$Country` , log10(ratio*1000)), y=log10(ratio*1000))) +
  geom_col() +
  scale_y_continuous(labels = function(x) 10^x/100) +
  labs(x = ' ',
       y= "Ratio of seized rounds of ammunition to seized arms
(logarithmic scale)") +
  ggtitle("Seizures of ammunition in relation to seizures of arms, by country, 2017") +
  theme_minimal() +
  geom_hline(yintercept = log10(mediana*100), linetype = "dashed", color = "red") +
  geom_hline(yintercept = log10(100), linetype = "dashed", color = "red")   +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = log10(cuartil_ratio[1]*100), ymax = log10(cuartil_ratio[2]*100)),
            fill = "red", alpha = 0.003) 

```

Next we will change the format of the text displayed on the X and Y
axes. Similarly, we rotate the names on the X-axis by 90º, in order to
make our graphics clearer.

```{r echo=TRUE}

  ggplot(ratiocalc, aes(x=reorder(`data$Country` , log10(ratio*1000)), y=log10(ratio*1000))) +
    geom_col() +
    scale_y_continuous(labels = function(x) 10^x/100) +
    labs(x = ' ',
         y= "Ratio of seized rounds of ammunition to seized arms
(logarithmic scale)") +
    ggtitle("Seizures of ammunition in relation to seizures of arms, by country, 2017") +
    theme_minimal() +
    scale_fill_manual(values = c("blue")) +
    theme(axis.text.x = element_text(face = "bold", size = 8, vjust = -0.5)) +
    theme(axis.text.y = element_text(face = "bold", size = 14)) +
    geom_hline(yintercept = log10(mediana*100), linetype = "dashed", color = "red") +
    geom_hline(yintercept = log10(100), linetype = "dashed", color = "red") +
    annotate("segment", x = 0, y = log10(mediana*100), xend = 0.1, yend = log10(mediana*100), color = "red")   +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = log10(cuartil_ratio[1]*100), ymax = log10(cuartil_ratio[2]*100)),
            fill = "red", alpha = 0.003)  +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

```

Now, we can proceed to the introduction of aesthetic elements related to
colour.

In this section we are interested in modifying: The background colour,
the colour of the bars and the colour of the line relative to the ratio
= 1.

In the same way, we will proceed in framing the graph:

```{r echo=TRUE}

  ggplot(ratiocalc, aes(x=reorder(`data$Country` , log10(ratio*1000)), y=log10(ratio*1000))) +
  geom_col(width = 0.5, fill = "#D2691E", position = "identity") +
  scale_y_continuous(labels = function(x) 10^x/100) +
  labs(x = ' ',
       y= "Ratio of seized rounds of ammunition to seized arms
(logarithmic scale)") +
  ggtitle("Seizures of ammunition in relation to seizures of arms, by country, 2017") +
  theme_minimal() +
  scale_fill_manual(values = c("blue")) +
  theme(axis.text.x = element_text(face = "bold", size = 8, vjust = -0.5)) +
  theme(axis.text.y = element_text(face = "bold", size = 14)) +
  geom_hline(yintercept = log10(mediana*100), linetype = "solid", size = 1, color = "red") +
  geom_hline(yintercept = log10(100), linetype = "solid", size = 1, color = "black") +
  annotate("segment", x = 0, y = log10(mediana*100), xend = 0.1, yend = log10(mediana*100), color = "red")   +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = log10(cuartil_ratio[1]*100), ymax = log10(cuartil_ratio[2]*100)),
            fill = "red", alpha = 0.003)  +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(panel.background = element_rect(fill = "lightgrey"),
        plot.background = element_rect(fill = "white"))

```

Next, we remove the grid from the bottom of the graph, leaving us with
our final replica.

### Final graph

```{r echo=TRUE, preview=TRUE}

ggplot(ratiocalc, aes(x=reorder(`data$Country` , log10(ratio*1000)), y=log10(ratio*1000))) +
  geom_col(width = 0.5, fill = "#D2691E", position = "identity") +
  scale_y_continuous(labels = function(x) 10^x/100) +
  labs(x = ' ',
       y= "Ratio of seized rounds of ammunition to seized arms
(logarithmic scale)") +
  ggtitle("Seizures of ammunition in relation to seizures of arms, by country, 2017") +
  theme_minimal() +
  scale_fill_manual(values = c("blue")) +
  theme(axis.text.x = element_text(face = "bold", size = 8, vjust = -0.5)) +
  theme(axis.text.y = element_text(face = "bold", size = 14)) +
  geom_hline(yintercept = log10(mediana*100), linetype = "solid", size = 1, color = "red") +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = log10(cuartil_ratio[1]*100), ymax = log10(cuartil_ratio[2]*100)),
            fill = "red", alpha = 0.003) +
  annotate("segment", x = -0.1, y = log10(100), xend = -0.1, yend = log10(1000), linetype = "solid") +
  geom_hline(yintercept = log10(100), linetype = "solid", size = 1, color = "black") +
  annotate("segment", x = 0, y = log10(mediana*100), xend = 0.1, yend = log10(mediana*100), color = "red") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(panel.background = element_rect(fill = "#EEE9E9"),
        plot.background = element_rect(fill = "white")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

```

### Limitations

While this graph represents an extremely clear identification of those
countries that are on either side of the ratio = 1, this information is
almost useless for gaining a regional (or global) understanding of the
relationships between countries where arms trafficking occurs, their
neighbours, and the regional circumstances that may cause this (regional
conflicts, paramilitary groups, etc.).

It is for this reason that I consider that the best possible improvement
to be made to this graph would be to transfer its data to a world map.

## Improvements

To convert the information of our dataset we must use the functions
related to the 'rnaturalearth' library, creating variables that
represent in R our world map.

Subsequently, we will establish relationships between the variables
'Country' and the variable 'Ratio', since they are the ones we want to
represent.

```{r echo=TRUE}

world <- ne_countries(scale = "medium", returnclass = "sf")

names(ratiocalc) <- c("Country", "ratio")

merged_data <- merge(world, ratiocalc, by.x = "name", by.y = "Country", all.x = TRUE)

```

## Final Map

Finally, we run the plot that will allow us to visualise our map:

```{r}

ggplot(merged_data, aes(fill = log10(ratio*1000))) +
  geom_sf() +
  scale_fill_distiller(palette = "RdYlBu", direction = 1) +  # Cambia el esquema de colores según tus preferencias
  labs(fill = "Ratio",
       title = "Seizures of ammunition in relation to seizures of arms, by country, 2017") +
  theme_minimal()


```

## Analysis and results

Although it is true that the database lacks the necessary information to
establish a global mapping, the extrapolation of the data present in our
dataset does allow us to visualise more clearly the disposition of
certain regional focal points.

### Meso-America and South-America

In the Mesoamerica and South America region we have data for: Guatemala,
Ecuador, Peru, Guyana, Brazil and Argentina.

Here Argentina has the lowest ratio in the region by far. The
interpretation of the data leads us to think that, at the regional
level, Argentina has the highest levels with respect to the number of
arms seized per quantity of ammunition seized. In this context, we could
situate the dynamics of illegal arms trafficking in Brazil as the most
diametrically opposed to those of Argentina. This would indicate that
more ammunition has been seized in Brazil than in Argentina. The rest of
the region is increasingly closer to a higher level of arms trafficking
as we move northwards (Peru, Ecuador, Guyana and Guatemala).

In the absence of data that could provide information on the behaviour
of the region as a whole, the overall reading is trivial, although it
may be far from reality.

The fact that Argentina is the country in the region most specialised in
arms trafficking could be due to its regional status. In this regard,
Argentina does not have any internal conflicts, so it is to be expected
that illegal arms trafficking is aimed at marketing arms to third
countries.

The violence in the Southern Macrozone in Chile or the narco-terrorist
insurgencies in Peru, Paraguay, Colombia and Ecuador can more than
justify the need to acquire military materiel. Similarly, the fact that
the data is so disproportionately in favour of Argentina would support
the idea of the latter as a 'proxy state' on which a regional network of
illegal arms trafficking could be built.

### Mediterrenean Sea

Regarding the Mediterranean, we have data from the following countries:
Morocco, Algeria, Tunisia, Libya and Spain.

The data with respect to Spain, which indicates more arms trafficking
than ammunition trafficking, could be explained by its geographical
position as the 'gateway to the Mediterranean' and a maritime exit to
the North Atlantic. Further research with data from continental Europe
could shed more light on this.

On the other side of the Strait of Gibraltar, Libya is the only country
that shows similar figures to Spain. The Second Libyan Civil War (2014 -
) left the country fractured into different centres of power with access
to different natural resources of vital importance for the functioning
of the country as a whole. In a state where power is shared between
local militias, terrorist groups, internationally mandated governments
and unrecognised institutions, armed conflict is already an entrenched
problem.

This would explain why illegal arms trafficking in Libya has a greater
tendency towards arms, but maintains a certain level of presence in
ammunition.

As for Algeria, the large quantity of seized ammunition could be due to
the surplus of military materiel from the colonial era. While Morocco
underwent a similar colonial process, the extent to which France armed
its former colony is greater than that which occurred on Moroccan
territory. The high levels of corruption in Morocco could explain a weak
military chain of command, which sought to take advantage of the
surplus. In a context of direct rivalry with Morocco (and with third
states involved), it is not at all inconceivable that Algeria would
consider selling military surplus to third countries as a means of
financing.

With respect to the rest of the regions, the dataset provides
insufficient information to make any kind of reading in this regard.

## Further research

The motivation for this work arose from the intention to identify
regional trafficking dynamics around existing conflicts. The next step
(once identified) would be to give them a reason.

I believe that the best way to address the issue of illegal arms
trafficking is to situate national contexts in the regional setting.
These data (and their visualisation) gave us the opportunity to see
(through their visualisation) different dynamics taking place in
different regional settings.

I believe that this work, accompanied by historical, political and
economic contextualisation, would reveal in depth the regional trends of
illegal arms trafficking and their relationship to the geopolitical
context in which it takes place.

From the numerical and historical analysis of regional trends in illicit
arms trafficking we can expect a clear motivation behind any illegal
transaction. The same motivation can give rise to a certain modus
operandi, refined over time. Identifying such patterns and associating
them with patterns would help to map (at a higher level) all the
elements and actors involved in illegal arms trafficking. This would
help to prevent and eradicate it.
