---
title: "Conflict in Sub-Suharan Africa"
description: |
  A graph representing conflict in 2021 - 2022.
categories: "2025"
author: Benjamin Sam
date: 2025-12-16
output:
  distill::distill_article:
    self_contained: false
    toc: true
---

## PACKAGES
```{r}
library(giscoR)
library(sf)
library(ggplot2)
library(ggforce)
library(patchwork)
library(dplyr)
library(grid)
library(png)
library(tidyverse)
library(lubridate)
```

## CLEANING THE DATA
```{r}
data <- read_csv("data/Africa_aggregated_data_up_to-2025-11-29.csv")

data <- 
  data |> 
  mutate(WEEK = dmy(WEEK))

filtered_data <- 
  data |> 
  filter(WEEK >= as.Date("2021-03-01") & WEEK <= as.Date("2022-04-30"))

country_list <- c("Mali", "Burkina Faso", "Niger", "Chad", 
                  "Sudan", "Nigeria", "Cameroon", "Central African Republic", 
                  "South Sudan", "Ethiopia", "Somalia", 
                  "Democratic Republic of Congo", "Uganda", "Mozambique")

filtered_data <- 
  filtered_data |>
  filter(COUNTRY %in% country_list) |> 
  group_by(COUNTRY) |> 
  summarise(total_fatalities = sum(FATALITIES))

# lookup table
values_lookup <- 
  tribble(
    ~COUNTRY, ~human_impact, ~incidence, ~geopolitical_impact,
    "Burkina Faso", 10, 15, 17,
    "Cameroon", 6, 7, 7,
    "Central African Republic", 8, 5, 44,
    "Chad", 2, 1, 23,
    "Democratic Republic of Congo", 32, 25, 42,
    "Ethiopia", 24, 13, 8,
    "Mali", 5, 10, 48,
    "Mozambique", 4, 4, 24,
    "Niger", 2, 3, 33,
    "Nigeria", 26, 27, 12,
    "Somalia", 20, 24, 42,
    "South Sudan", 19, 7, 45,
    "Sudan",19, 6, 30,
    "Uganda", 0, 3, 7
)

filtered_data <- 
  filtered_data |> 
  left_join(values_lookup, by = "COUNTRY")
```

## BASE MAP
### COLOURS FOR ARC GRAPHS
```{r}
color_points <- tibble(
  value = c(0, 5, 10, 15, 17, 20, 23, 30, 32, 42, 44, 48),
  color = c("#FAF8B2", "#FAF3B0", "#F6E5A8", "#F0DA9F", "#EFD69C",
            "#EDCF97", "#E9C790", "#E4B585", "#E3B080", "#D89670",
            "#D9956D", "#CF8A69")
) %>%
  arrange(value)

value_to_color <- function(val) {
  if (val %in% color_points$value) {
    return(color_points$color[color_points$value == val])
  } else {
    lower_idx <- max(which(color_points$value < val))
    upper_idx <- min(which(color_points$value > val))
    
    lower_val <- color_points$value[lower_idx]
    upper_val <- color_points$value[upper_idx]
    lower_col <- color_points$color[lower_idx]
    upper_col <- color_points$color[upper_idx]
    
    weight <- (val - lower_val) / (upper_val - lower_val)
    return(colorRamp(c(lower_col, upper_col))(weight) %>%
             rgb(maxColorValue = 255))
  }
}

# Get African countries from giscoR
countries <- gisco_get_countries(
  year = "2020",
  epsg = "4326",
  resolution = "20"
)

# Filter to region
bbox <- st_bbox(c(xmin = -27, xmax = 64, ymin = -31, ymax = 41),
                crs = st_crs(4326))
bbox_poly <- st_as_sfc(bbox)
countries_in_view <- countries[st_intersects(countries, bbox_poly, sparse = FALSE), ]

# Define country colors
# Orange: conflict-affected countries
# White: other Sub-Saharan countries
# Beige: Other countries on the map
country_colors <- tribble(
  ~COUNTRY, ~color,
  # Conflict-affected (orange-y)
  "Mali", "#ED8758",
  "Burkina Faso", "#EC8657",
  "Niger", "#F29464",
  "Chad", "#F7B88D",
  "Sudan", "#F18B5C",
  "Abyei", "#F08A5B",
  "South Sudan", "#F08A5B",
  "Ethiopia", "#DD6F4B",
  "Somalia", "#ED8556",
  "Uganda", "#F6D1A5",
  "Democratic Republic of The Congo", "#E67D52",
  "Mozambique", "#F19F70",
  "Nigeria", "#D46443",
  "Cameroon", "#F3A172",
  "Central African Republic", "#F28C5D",
  # Other Sub-Saharan (white)
  "Mauritania", "#FFFFFF",
  "Senegal", "#FFFFFF",
  "Gambia", "#FFFFFF",
  "Guinea-Bissau", "#FFFFFF",
  "Cape Verde", "#FFFFFF",
  "Guinea", "#FFFFFF",
  "Sierra Leone", "#FFFFFF",
  "Côte D’Ivoire", "#FFFFFF",
  "Ghana", "#FFFFFF",
  "Togo", "#FFFFFF",
  "Benin", "#FFFFFF",
  "Gabon", "#FFFFFF",
  "Congo", "#FFFFFF",
  "Angola", "#FFFFFF",
  "Zambia", "#FFFFFF",
  "Namibia", "#FFFFFF",
  "Botswana", "#FFFFFF",
  "Zimbabwe", "#FFFFFF",
  "South Africa", "#FFFFFF",
  "Madagascar", "#FFFFFF",
  "Malawi", "#FFFFFF",
  "United Republic of Tanzania", "#FFFFFF",
  "Burundi", "#FFFFFF",
  "Rwanda", "#FFFFFF",
  "Kenya", "#FFFFFF",
  "Eritrea", "#FFFFFF",
  "Djibouti", "#FFFFFF",
  "Liberia", "#FFFFFF",
  "Equatorial Guinea", "#FFFFFF",
  "Eswatini", "#FFFFFF",
  "Lesotho", "#FFFFFF",
)

# Merge colors with map data
countries_in_view_colored <- countries_in_view %>%
  left_join(country_colors, by = c("NAME_ENGL" = "COUNTRY")) %>%
  mutate(fill_color = ifelse(is.na(color), "#F0EDE3", color))

# Create coastline
all_countries_union <- st_union(countries_in_view)
coastline <- st_cast(st_boundary(all_countries_union), "LINESTRING")

# COUNTRY LABELS
# Auto-labeled countries (using centroids)
auto_countries <- c(
  "Niger", 
  "Chad", 
  "Senegal",  
  "Liberia", 
  "Togo", 
  "Gabon", 
  "Angola", 
  "Namibia",
  "South Africa", 
  "Botswana", 
  "Zimbabwe",  
  "Malawi",
  "Madagascar", 
  "Burundi", 
  "Rwanda", 
  "Kenya",
  "Djibouti"
)

countries_auto <- countries_in_view |>
  filter(NAME_ENGL %in% auto_countries)

# Precisely positioned countries (many did not behave well. Here the text at the base of the plot about the sources of the data was plotted as well)
custom_positions <- tribble(
  ~name, ~lon, ~lat,
    "SOUTH\nSUDAN", 30, 8.0,
  "CENTRAL AFRICAN\nREPUBLIC", 20.0, 6.5,
  "SOMALIA", 45.5, 4.2,
  "ETHIOPIA", 40.5, 9.0,
  "BURKINA\nFASO", -0.5, 12.5,
  "MOZAMBIQUE", 35.0, -18.0,
  "UGANDA", 33, 2.5,
  "SUDAN", 30.0, 15.0,
  "CABO\nVERDE", -23.5, 18.5,
  "MALI", -7, 14,
  "ERITREA", 38, 17,
  "TANZANIA", 35, -6,
  "CÔTE\nD'IVOIRE", -5.2, 8,
  "DEMOCRATIC\nREPUBLIC\nOF THE CONGO", 23.5, -5,
  "GUINEA-\nBISSAU", -18, 11,
  "REPUBLIC\nOF CONGO", 16, 1,
  "EQUATORIAL\nGUINEA", 5.5, 2,
  "CAMEROON", 12, 5,
  "BENIN", 2.5, 10.5,
  "GHANA", -1.5, 6.5,
  "SIERRA\nLEONE", -11.5, 9,
  "GUINEA", -11, 11,
  "THE GAMBIA", -18, 13,
  "MAURITANIA", -13, 19.5,
  "NIGERIA", 6.5, 9.5,
  "ZAMBIA", 25, -15,
  "Sources: Armed Conflict Location & Event Data Project (ACLED), www.acleddata.com; IISS analysis", -6, -29
) |>
  st_as_sf(coords = c("lon", "lat"), crs = 4326)

# FINAL MAP
# Incorporates countries with colours, the coastline, the labels and the ocean colour
base_map <- ggplot() +
  geom_sf(data = countries_in_view_colored,
          aes(fill = fill_color),
          color = "#B8C0BE",
          linewidth = 0.2) +
  scale_fill_identity() +
  geom_sf(data = coastline, color = "#387298", linewidth = 0.2) +
  geom_sf_text(data = countries_auto,
               aes(label = toupper(NAME_ENGL)),
               fontface = "plain", 
               size = 2.3, 
               lineheight = 0.9) +
  geom_sf_text(data = custom_positions,
               aes(label = name),
               fontface = "plain",
               size = 2.3,
               lineheight = 0.9) +
  coord_sf(xlim = c(-27, 64), ylim = c(-31, 41), expand = FALSE) +
  theme_void() +
  theme(panel.background = element_rect(fill = "#DCE8F2"),
        plot.background = element_rect(fill = "white"))
```


### UPDATED DATA CREATION FOR COUNTRY GRAPHS
- Function to create country data
- I found myself repeating the same lines of code many, many times, so I used Claude to help me build a function to avoid the repetition
```{r}
create_country_data <- function(country_name, filtered_data, value_to_color) {
  filtered_data |> 
    filter(COUNTRY == country_name) |> 
    pivot_longer(
      cols = c(human_impact, incidence, geopolitical_impact),
      names_to = "category",
      values_to = "value"
    ) |> 
    mutate(original_segment = row_number()) |> 
    arrange(desc(row_number())) |> 
    mutate(
      color = map_chr(value, value_to_color),
      segment = row_number(),
      original_start = pi - ((original_segment - 1) / 3 * pi),
      original_end = pi - (original_segment / 3 * pi),
      start_angle = (pi - ((segment - 1) / 3 * pi)) - pi/2,
      end_angle = (pi - (segment / 3 * pi)) - pi/2,
      mid_angle = (original_start + original_end) / 2,
      label_x = cos(mid_angle) * 1.08,
      label_y = sin(mid_angle) * 1.08
    )
}

# Create all country data frames using the function
cameroon_data <- create_country_data("Cameroon", filtered_data, value_to_color)
central_african_republic_data <- create_country_data("Central African Republic", filtered_data, value_to_color)
chad_data <- create_country_data("Chad", filtered_data, value_to_color)
democratic_republic_of_congo_data <- create_country_data("Democratic Republic of Congo", filtered_data, value_to_color)
ethiopia_data <- create_country_data("Ethiopia", filtered_data, value_to_color)
mali_data <- create_country_data("Mali", filtered_data, value_to_color)
mozambique_data <- create_country_data("Mozambique", filtered_data, value_to_color)
niger_data <- create_country_data("Niger", filtered_data, value_to_color)
nigeria_data <- create_country_data("Nigeria", filtered_data, value_to_color)
somalia_data <- create_country_data("Somalia", filtered_data, value_to_color)
south_sudan_data <- create_country_data("South Sudan", filtered_data, value_to_color)
sudan_data <- create_country_data("Sudan", filtered_data, value_to_color)
uganda_data <- create_country_data("Uganda", filtered_data, value_to_color)
burkina_faso_data <- create_country_data("Burkina Faso", filtered_data, value_to_color)
```
## COUNTRY ARC PLOTS

- Function to create arc plots
-  These proved unexpectedly challenging. There was a package that worked when I began the project to plot parliament graphs, however this stopped working midway through writing the project. I resorted to stack exchange and Claude to help me build the graphs.
```{r}
create_arc_plot <- function(data, country_name, is_tall = FALSE) {
  ymin_val <- if(is_tall) -0.8 else -0.6
  ylim_val <- if(is_tall) c(-0.8, 1.0) else c(-0.6, 1.0)
  text_y <- if(is_tall) -0.55 else -0.42
  
  # Create base plot
  p <- ggplot(data) +
    annotate("rect", xmin = -1.2, xmax = 1.2, ymin = ymin_val, ymax = 1,
             fill = alpha("white", 0.5), color = "black", linewidth = 0.2) +
    geom_arc_bar(aes(x0 = 0, y0 = -0.3, r0 = 0.6, r = 0.9,
                     start = start_angle, end = end_angle, fill = color),
                 color = "white", linewidth = 0.2) +
    scale_fill_identity()
  
  # Add text layers conditionally based on the height of the plot (the placement of the text differs for 2 of the plots as the names of the countries are longer, so they needed to be taller.)
  
  if (is_tall) {
    p <- p +
      geom_text(aes(x = label_x * 1.05, y = (label_y - 0.3) * 1.05, label = value),
                fontface = "plain", size = 2.75, lineheight = 0.9) +
      annotate("text", x = 0, y = text_y, label = country_name, 
               fontface = "bold", size = 2.75, lineheight = 0.9)
  } else {
    p <- p +
      geom_text(aes(x = label_x * 1.05, y = (label_y - 0.3) * 1.05, label = value),
                fontface = "plain", size = 2.75, lineheight = 0.9) +
      annotate("text", x = 0, y = text_y, label = country_name, 
               fontface = "bold", size = 2.75, lineheight = 0.9)
  }
  
  p <- p +
    coord_fixed(xlim = c(-1.3, 1.3), ylim = ylim_val) +
    theme_void() +
    theme(plot.background = element_rect(fill = NA, color = NA),
          plot.margin = margin(0, 0, 0, 0),
          text = element_text(size = rel(1)))
  
  return(p)
}

# Create all arc plots
cameroon_plot <- create_arc_plot(cameroon_data, "Cameroon")
central_african_republic_plot <- create_arc_plot(central_african_republic_data, "Central African\nRepublic", is_tall = TRUE)
chad_plot <- create_arc_plot(chad_data, "Chad")
democratic_republic_of_congo_plot <- create_arc_plot(democratic_republic_of_congo_data, "Democratic Rep.\nof the Congo", is_tall = TRUE)
ethiopia_plot <- create_arc_plot(ethiopia_data, "Ethiopia")
mali_plot <- create_arc_plot(mali_data, "Mali")
mozambique_plot <- create_arc_plot(mozambique_data, "Mozambique")
niger_plot <- create_arc_plot(niger_data, "Niger")
nigeria_plot <- create_arc_plot(nigeria_data, "Nigeria")
somalia_plot <- create_arc_plot(somalia_data, "Somalia")
south_sudan_plot <- create_arc_plot(south_sudan_data, "South Sudan")
sudan_plot <- create_arc_plot(sudan_data, "Sudan")
uganda_plot <- create_arc_plot(uganda_data, "Uganda")
burkina_faso_plot <- create_arc_plot(burkina_faso_data, "Burkina Faso")
```

### BASE MAP + LINES FROM ARC PLOTS
- Helper function to convert pixels to lon/lat coordinates
- This let me measure the positions of the lines as a number of pixels (in preview on mac this appears when you drag a selection box on an image) and have that series of "pixel co-ordinates" converted to latitude and longitude to plot on the map.
```{r}
pixel_to_lonlat <- function(x_pixels, y_pixels, 
                            map_width = 2232, map_height = 1783,
                            lon_min = -27, lon_max = 64,
                            lat_min = -31, lat_max = 41) {
  x_prop <- x_pixels / map_width
  y_prop <- y_pixels / map_height
  lon <- lon_min + x_prop * (lon_max - lon_min)
  lat <- lat_max - y_prop * (lat_max - lat_min)
  return(list(lon = lon, lat = lat))
}

# ALL CONNECTION LINES (these are the pixel measurements discussed above. You read them as pairs; distance from the keft edge of the map is the x and from the top is the y. Taken together they define points for the lines to be plotted to.)
connections <- list(
  #arc graphs
  mali = list(
    x = c(298, 298, 445, 445),
    y = c(212, 250, 250, 670)
  ),
  burkina_faso = list(
    x = c(586, 586),
    y = c(209, 714)
  ),
  niger = list(
    x = c(869, 869),
    y = c(209, 531)
  ),
  chad = list(
    x = c(1131, 1131),
    y = c(209, 600)
  ),
  sudan = list(
    x = c(1379, 1379),
    y = c(209, 576)
  ),
  south_sudan = list(
    x = c(1630, 1630, 1459, 1459),
    y = c(330, 519, 519, 886)
  ),
  ethiopia = list(
    x = c(1967, 1612, 1612),
    y = c(627, 627, 768)
  ),
  somalia = list(
    x = c(1967, 1871, 1871, 1800),
    y = c(810, 810, 887, 887)
  ),
  uganda = list(
    x = c(1967, 1450),
    y = c(992, 992)
  ),
  drc = list(
    x = c(1967, 1351),
    y = c(1185, 1185)
  ),
  mozambique = list(
    x = c(1967, 1477),
    y = c(1593, 1593)
  ),
  central_african_republic = list(
    x = c(1090, 1126, 1126, 1090),
    y = c(1197, 1197, 885, 885)
  ),
  cameroon = list(
    x = c(557, 874, 874, 955, 955),
    y = c(1051, 1051, 940, 940, 921)
  ),
  nigeria = list(
    x = c(308, 840, 840),
    y = c(924, 924, 807)
  ),
  #country/flag boxes
  the_sahel_left = list(
    x = c(613, 613),
    y = c(553, 600)
  ),
  the_sahel_middle = list(
    x = c(644, 644),
    y = c(553, 661)
  ),
  the_sahel_right = list(
    x = c(804, 804),
    y = c(553, 605)
  ),
  lake_chad_basin_top_left = list(
    x = c(975, 975, 882),
    y = c(664, 609, 609)
  ),
  lake_chad_basin_bottom_left = list(
    x = c(902, 902, 862),
    y = c(790, 822, 822)
  ),
  lake_chad_basin_bottom = list(
    x = c(971, 971),
    y = c(790, 841)
  ),
  lake_chad_basin_right = list(
    x = c(1129, 1157, 1157),
    y = c(721, 721, 662)
  ),
  great_lakes_region_bottom = list(
    x = c(1252, 1252),
    y = c(1061, 1084)
  ),
  great_lakes_region_right = list(
    x = c(1340, 1447),
    y = c(1023, 1023)
  ),
  rawanda_eu_sadcmm = list(
    x = c(1697, 1697, 1550),
    y = c(1370, 1437, 1437)
  ),
  turkey_uk_us_au_eu = list(
    x = c(1773, 1773),
    y = c(1022, 942)
  ),
  eritrea = list(
    x = c(1657, 1681),
    y = c(862, 862)
  ),
  un_top = list(
    x = c(1332, 1332),
    y = c(687, 653)
  ),
  un_bottom = list(
    x = c(1333, 1333),
    y = c(750, 850)
  ),
  russia_rwanda_un = list(
    x = c(1252, 1252, 1191),
    y = c(445, 785, 785)
  )
)

# white endpoints
# The lines used for the arc graphs and countries involved in the various conflicts are differentiated by the colour of the end points they use.
white_endpoints <- c("the_sahel_left", 
                     "the_sahel_middle", 
                     "the_sahel_right", 
                     "lake_chad_basin_top_left", 
                     "lake_chad_basin_bottom_left", 
                     "lake_chad_basin_bottom",
                     "lake_chad_basin_right",
                     "great_lakes_region_bottom",
                     "great_lakes_region_right",
                     "rawanda_eu_sadcmm_bottom",
                     "rawanda_eu_sadcmm",
                     "turkey_uk_us_au_eu",
                     "eritrea",
                     "un_top",
                     "un_bottom",
                     "russia_rwanda_un")

# Lines
lines_df <- map_dfr(names(connections), function(country) {
  coords <- pixel_to_lonlat(
    x_pixels = connections[[country]]$x,
    y_pixels = connections[[country]]$y
  )
  data.frame(
    x = coords$lon,
    y = coords$lat,
    country = country
  )
})

# Endpoints
# This section makes a dataframe using the final coordinates of the lines as the position of the endpoint.
endpoints_df <- map_dfr(names(connections), function(country) {
  n <- length(connections[[country]]$x)
  coords <- pixel_to_lonlat(
    x_pixels = connections[[country]]$x[n],
    y_pixels = connections[[country]]$y[n]
  )
  data.frame(
    x = coords$lon, 
    y = coords$lat, 
    country = country,
    is_white = country %in% white_endpoints
  )
})

# Split into black and white endpoints
black_endpoints <- endpoints_df |> filter(!is_white)
white_endpoints_df <- endpoints_df |> filter(is_white)

# Add lines and endpoints to map
base_map <- base_map +
  geom_path(data = lines_df,
            aes(x = x, y = y, group = country),
            color = "black",
            linewidth = 0.3) +
  # Black endpoints (fill and color both black)
  geom_point(data = black_endpoints,
             aes(x = x, y = y),
             color = "black",
             fill = "black",
             shape = 21,
             size = 1.5,
             stroke = 0.3) +
  # White endpoints
  geom_point(data = white_endpoints_df,
             aes(x = x, y = y),
             color = "black",
             fill = "white",
             shape = 21,
             size = 1.5,
             stroke = 0.3)
```
### BASE MAP + ARC GRAPHS
```{r}
# Standard arc box dimensions
# These dimensions applied for the majority of the arc graphs
arc_w <- 215 / 2232
arc_h <- 156 / 1783

# CAR and DRC dimensions
# I realised quite late that these graphs were subtly different in height due to the length of the country name needing to fit in the box
car_drc_h <- 183 / 1783

# Map dimensions
map_width <- 2232
map_height <- 1783

# Helper function for positioning
# I reused the same method as for the lines, this time measuring to the left edge of the plots and the bottom edge.
pixel_to_prop <- function(left_px, top_px, height_px = 156) {
  list(
    left = left_px / map_width,
    bottom = 1 - (top_px + height_px) / map_height
  )
}

# Helper function to add inset to map
add_inset <- function(map, plot, left, bottom, width, height) {
  map + inset_element(
    plot,
    left = left,
    right = left + width,
    bottom = bottom,
    top = bottom + height
  )
}

# Define all arc plot configurations
arc_configs <- list(
  list(plot = mali_plot, pos = pixel_to_prop(190, 77), w = arc_w, h = arc_h),
  list(plot = burkina_faso_plot, pos = pixel_to_prop(487, 77), w = arc_w, h = arc_h),
  list(plot = niger_plot, pos = pixel_to_prop(754, 77), w = arc_w, h = arc_h),
  list(plot = chad_plot, pos = pixel_to_prop(1020, 77), w = arc_w, h = arc_h),
  list(plot = sudan_plot, pos = pixel_to_prop(1272, 77), w = arc_w, h = arc_h),
  list(plot = south_sudan_plot, pos = pixel_to_prop(1528, 196), w = arc_w, h = arc_h),
  list(plot = ethiopia_plot, pos = pixel_to_prop(1947, 547), w = arc_w, h = arc_h),
  list(plot = somalia_plot, pos = pixel_to_prop(1947, 733), w = arc_w, h = arc_h),
  list(plot = uganda_plot, pos = pixel_to_prop(1947, 914), w = arc_w, h = arc_h),
  list(plot = democratic_republic_of_congo_plot, pos = pixel_to_prop(1947, 1108, height_px = 183), w = arc_w, h = car_drc_h),
  list(plot = mozambique_plot, pos = pixel_to_prop(1947, 1517), w = arc_w, h = arc_h),
  list(plot = nigeria_plot, pos = pixel_to_prop(114, 836), w = arc_w, h = arc_h),
  list(plot = cameroon_plot, pos = pixel_to_prop(364, 943), w = arc_w, h = arc_h),
  list(plot = central_african_republic_plot, pos = pixel_to_prop(897, 1104, height_px = 183), w = arc_w, h = car_drc_h)
)

# Add all arc plots to map
for (config in arc_configs) {
  base_map <- add_inset(base_map, config$plot, config$pos$left, config$pos$bottom, config$w, config$h)
}
```
## BOXES
### SAHEL BOX
- I used a different approach here as there were a lot of elements to position. I measured the flags and overall box size in pixels and then positioned them relative to the left and bottom edges of the boxes. This was fiddly, so I ended up measuring the margins, flag width, gap between the flags and plotting these for each box (you can see this in the code for the other boxes below)
```{r}
create_sahel_box <- function() {
  france_flag <- readPNG("flags_and_logos/france.png")
  italy_flag <- readPNG("flags_and_logos/italy.png")
  usa_flag <- readPNG("flags_and_logos/us.png")
  eu_logo <- readPNG("flags_and_logos/eu.png")
  un_logo <- readPNG("flags_and_logos/un.png")
  au_logo <- readPNG("flags_and_logos/au.png")
  g5_logo <- readPNG("flags_and_logos/sahel_g5.png")
  
  flag_width <- 65 / 245
  flag_height <- 43 / 205
  logo_width <- 65 / 245
  logo_height <- 43 / 205
  logo_circle_width <- 43 / 245
  logo_circle_height <- 43 / 205
  
  p <- ggplot() +
    annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1,
             fill = "white", color = NA) +
    annotate("text", x = 0.5, y = 0.92, label = "The Sahel",
             hjust = 0.5, fontface = "bold", size = 2.75, lineheight = 0.9) +
    annotation_raster(france_flag, 
                      xmin = 0.08, xmax = 0.08 + flag_width, 
                      ymin = 0.62, ymax = 0.62 + flag_height) +
    annotation_raster(italy_flag, 
                      xmin = 0.5 - (flag_width / 2), xmax = 0.5 + (flag_width / 2), 
                      ymin = 0.62, ymax = 0.62 + flag_height) +
    annotation_raster(usa_flag, 
                      xmin = 0.92 - flag_width, xmax = 0.92, 
                      ymin = 0.62, ymax = 0.62 + flag_height) +
    annotation_raster(eu_logo, 
                      xmin = 0.08, xmax = 0.08 + logo_width, 
                      ymin = 0.30, ymax = 0.30 + logo_height) +
    annotation_raster(un_logo, 
                      xmin = 0.5 - (logo_width / 2), xmax = 0.5 + (logo_width / 2), 
                      ymin = 0.30, ymax = 0.30 + logo_height) +
    annotation_raster(au_logo, 
                      xmin = 0.92 - logo_width, xmax = 0.92, 
                      ymin = 0.30, ymax = 0.30 + logo_height) +
    annotation_raster(g5_logo, 
                      xmin = 0.5 - (logo_circle_width / 2), 
                      xmax = 0.5 + (logo_circle_width / 2), 
                      ymin = 0.02, 
                      ymax = 0.02 + logo_circle_height) +
    coord_fixed(ratio = 205/245, xlim = c(0, 1), ylim = c(0, 1)) +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "white", color = "black", 
                                     linewidth = 0.5, radius = unit(0, "pt")),
      plot.margin = margin(0, 0, 0, 0)
    )
  
  return(p)
}
```
### LAKE CHAD BOX
```{r}
create_lake_chad_box <- function() {
  france_flag <- readPNG("flags_and_logos/france.png")
  uk_flag <- readPNG("flags_and_logos/uk.png")
  usa_flag <- readPNG("flags_and_logos/us.png")
  multi_jtf_logo <- readPNG("flags_and_logos/multi.png")
  
  left_margin <- 50 / 235
  flag_width <- 65 / 235
  gap <- 5 / 235
  flag_height <- 43 / 145
  logo_width <- 65 / 235
  logo_height <- 43 / 145
  
  p <- ggplot() +
    annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1,
             fill = "white", color = NA) +
    annotate("text", x = 0.5, y = 0.88, label = "Lake Chad Basin",
             hjust = 0.5, fontface = "bold", size = 2.75, lineheight = 0.9) +
    annotation_raster(france_flag, 
                      xmin = left_margin, 
                      xmax = left_margin + flag_width, 
                      ymin = 0.4, 
                      ymax = 0.4 + flag_height) +
    annotation_raster(uk_flag, 
                      xmin = left_margin + flag_width + gap, 
                      xmax = left_margin + flag_width + gap + flag_width, 
                      ymin = 0.4, 
                      ymax = 0.4 + flag_height) +
    annotation_raster(usa_flag, 
                      xmin = left_margin, 
                      xmax = left_margin + flag_width, 
                      ymin = 0.06, 
                      ymax = 0.06 + flag_height) +
    annotation_raster(multi_jtf_logo, 
                      xmin = left_margin + flag_width + gap, 
                      xmax = left_margin + flag_width + gap + logo_width, 
                      ymin = 0.06, 
                      ymax = 0.06 + logo_height) +
    coord_fixed(ratio = 145/235, xlim = c(0, 1), ylim = c(0, 1)) +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "white", color = "black", 
                                     linewidth = 0.5, radius = unit(0, "pt")),
      plot.margin = margin(0, 0, 0, 0)
    )
  
  return(p)
}
```
### GREAT LAKES BOX
```{r}
create_great_lakes_box <- function() {
  un_logo <- readPNG("flags_and_logos/un.png")
  au_logo <- readPNG("flags_and_logos/au.png")
  rwanda_flag <- readPNG("flags_and_logos/rwanda.png")
  burundi_flag <- readPNG("flags_and_logos/burundi.png")
  
  box_width <- 177
  box_height <- 177
  left_margin <- 13 / box_width
  flag_width <- 65 / box_width
  gap <- 25 / box_width
  flag_height <- 43 / box_height
  logo_width <- 65 / box_width
  logo_height <- 43 / box_height
  
  p <- ggplot() +
    annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1,
             fill = "white", color = NA) +
    annotate("text", x = 0.5, y = 0.88, label = "Great\nLakes Region",
             hjust = 0.5, fontface = "bold", size = 2.75, lineheight = 0.9) +
    annotation_raster(au_logo, 
                      xmin = left_margin, 
                      xmax = left_margin + flag_width, 
                      ymin = 0.4, 
                      ymax = 0.4 + flag_height) +
    annotation_raster(un_logo, 
                      xmin = left_margin + flag_width + gap, 
                      xmax = left_margin + flag_width + gap + flag_width, 
                      ymin = 0.4, 
                      ymax = 0.4 + flag_height) +
    annotation_raster(rwanda_flag, 
                      xmin = left_margin, 
                      xmax = left_margin + flag_width, 
                      ymin = 0.06, 
                      ymax = 0.06 + flag_height) +
    annotation_raster(burundi_flag, 
                      xmin = left_margin + flag_width + gap, 
                      xmax = left_margin + flag_width + gap + logo_width, 
                      ymin = 0.06, 
                      ymax = 0.06 + logo_height) +
    coord_fixed(ratio = box_height/box_width, xlim = c(0, 1), ylim = c(0, 1)) +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "white", color = "black", 
                                     linewidth = 0.5, radius = unit(0, "pt")),
      plot.margin = margin(0, 0, 0, 0)
    )
  
  return(p)
}
```
### RWANDA EU SADCMM BOX
```{r}
create_two_row_box <- function() {
  rwanda_flag <- readPNG("flags_and_logos/rwanda.png")
  eu_logo <- readPNG("flags_and_logos/eu.png")
  sadcmm_flag <- readPNG("flags_and_logos/sadcmm.png")
  
  box_width <- 195
  box_height <- 135
  top_margin <- 65 / box_width
  top_flag_width <- 65 / box_width
  bottom_left_margin <- 22 / box_width
  bottom_flag_width <- 65 / box_width
  bottom_gap <- 21 / box_width
  flag_height <- 43 / box_height
  
  p <- ggplot() +
    annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1,
             fill = "white", color = NA) +
    annotation_raster(rwanda_flag, 
                      xmin = top_margin, 
                      xmax = top_margin + top_flag_width, 
                      ymin = 0.55, 
                      ymax = 0.55 + flag_height) +
    annotation_raster(eu_logo, 
                      xmin = bottom_left_margin, 
                      xmax = bottom_left_margin + bottom_flag_width, 
                      ymin = 0.15, 
                      ymax = 0.15 + flag_height) +
    annotation_raster(sadcmm_flag, 
                      xmin = bottom_left_margin + bottom_flag_width + bottom_gap, 
                      xmax = bottom_left_margin + bottom_flag_width + bottom_gap + bottom_flag_width, 
                      ymin = 0.15, 
                      ymax = 0.15 + flag_height) +
    coord_fixed(ratio = box_height/box_width, xlim = c(0, 1), ylim = c(0, 1)) +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "white", color = "black", 
                                     linewidth = 0.5, radius = unit(0, "pt")),
      plot.margin = margin(0, 0, 0, 0)
    )
  
  return(p)
}
```
### TURK UK US AU EU BOX
```{r}
create_three_two_box <- function() {
  top_flag1 <- readPNG("flags_and_logos/turkey.png")
  top_flag2 <- readPNG("flags_and_logos/uk.png")
  top_flag3 <- readPNG("flags_and_logos/us.png")
  bottom_flag1 <- readPNG("flags_and_logos/au.png")
  bottom_flag2 <- readPNG("flags_and_logos/eu.png")
  
  box_width <- 245
  box_height <- 138
  top_space <- 12.5 / box_width
  top_flag_width <- 65 / box_width
  bottom_flag_width <- 65 / box_width
  bottom_gap <- 10 / box_width
  bottom_margin <- 52.5 / box_width
  flag_height <- 43 / box_height
  
  p <- ggplot() +
    annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1,
             fill = "white", color = NA) +
    annotation_raster(top_flag1, 
                      xmin = top_space, 
                      xmax = top_space + top_flag_width, 
                      ymin = 0.55, 
                      ymax = 0.55 + flag_height) +
    annotation_raster(top_flag2, 
                      xmin = top_space * 2 + top_flag_width, 
                      xmax = top_space * 2 + top_flag_width * 2, 
                      ymin = 0.55, 
                      ymax = 0.55 + flag_height) +
    annotation_raster(top_flag3, 
                      xmin = top_space * 3 + top_flag_width * 2, 
                      xmax = top_space * 3 + top_flag_width * 3, 
                      ymin = 0.55, 
                      ymax = 0.55 + flag_height) +
    annotation_raster(bottom_flag1, 
                      xmin = bottom_margin, 
                      xmax = bottom_margin + bottom_flag_width, 
                      ymin = 0.15, 
                      ymax = 0.15 + flag_height) +
    annotation_raster(bottom_flag2, 
                      xmin = bottom_margin + bottom_flag_width + bottom_gap, 
                      xmax = bottom_margin + bottom_flag_width + bottom_gap + bottom_flag_width, 
                      ymin = 0.15, 
                      ymax = 0.15 + flag_height) +
    coord_fixed(ratio = box_height/box_width, xlim = c(0, 1), ylim = c(0, 1)) +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "white", color = "black", 
                                     linewidth = 0.5, radius = unit(0, "pt")),
      plot.margin = margin(0, 0, 0, 0)
    )
  
  return(p)
}
```
### ERITREA BOX
```{r}
create_single_flag_box <- function() {
  eritrea_flag <- readPNG("flags_and_logos/eritrea.png")
  
  box_width <- 95
  box_height <- 65
  flag_margin <- 15 / box_width
  flag_width <- 65 / box_width
  flag_height <- 43 / box_height
  flag_y_margin <- 11 / box_height
  
  p <- ggplot() +
    annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1,
             fill = "white", color = NA) +
    annotation_raster(eritrea_flag, 
                      xmin = flag_margin, 
                      xmax = flag_margin + flag_width, 
                      ymin = flag_y_margin, 
                      ymax = flag_y_margin + flag_height) +
    coord_fixed(ratio = box_height/box_width, xlim = c(0, 1), ylim = c(0, 1)) +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "white", color = "black", 
                                     linewidth = 0.5, radius = unit(0, "pt")),
      plot.margin = margin(0, 0, 0, 0)
    )
  
  return(p)
}
```
### UN BOX
```{r}
create_single_flag_box_2 <- function() {
  un_logo <- readPNG("flags_and_logos/un.png")
  
  box_width <- 95
  box_height <- 65
  flag_margin <- 15 / box_width
  flag_width <- 65 / box_width
  flag_height <- 43 / box_height
  flag_y_margin <- 11 / box_height
  
  p <- ggplot() +
    annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1,
             fill = "white", color = NA) +
    annotation_raster(un_logo, 
                      xmin = flag_margin, 
                      xmax = flag_margin + flag_width, 
                      ymin = flag_y_margin, 
                      ymax = flag_y_margin + flag_height) +
    coord_fixed(ratio = box_height/box_width, xlim = c(0, 1), ylim = c(0, 1)) +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "white", color = "black", 
                                     linewidth = 0.5, radius = unit(0, "pt")),
      plot.margin = margin(0, 0, 0, 0)
    )
  
  return(p)
}
```
### RUSSIA RWANDA UN BOX
```{r}
create_three_flag_row_box <- function() {
  russia_flag <- readPNG("flags_and_logos/russia.png")
  rwanda_flag <- readPNG("flags_and_logos/rwanda.png")
  un_logo <- readPNG("flags_and_logos/un.png")
  
  box_width <- 205
  box_height <- 84
  space <- 2.5 / box_width
  flag_width <- 65 / box_width
  flag_height <- 43 / box_height
  flag_y_margin <- (box_height - 43) / 2 / box_height
  
  p <- ggplot() +
    annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1,
             fill = "white", color = NA) +
    annotation_raster(russia_flag, 
                      xmin = space, 
                      xmax = space + flag_width, 
                      ymin = flag_y_margin, 
                      ymax = flag_y_margin + flag_height) +
    annotation_raster(rwanda_flag, 
                      xmin = space * 2 + flag_width, 
                      xmax = space * 2 + flag_width * 2, 
                      ymin = flag_y_margin, 
                      ymax = flag_y_margin + flag_height) +
    annotation_raster(un_logo, 
                      xmin = space * 3 + flag_width * 2, 
                      xmax = space * 3 + flag_width * 3, 
                      ymin = flag_y_margin, 
                      ymax = flag_y_margin + flag_height) +
    coord_fixed(ratio = box_height/box_width, xlim = c(0, 1), ylim = c(0, 1)) +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "white", color = "black", 
                                     linewidth = 0.5, radius = unit(0, "pt")),
      plot.margin = margin(0, 0, 0, 0)
    )
  
  return(p)
}
```
### CREATE ALL BOX PLOTS
```{r}
sahel_box <- create_sahel_box()
lake_chad_box <- create_lake_chad_box()
great_lakes_box <- create_great_lakes_box()
two_row_box <- create_two_row_box()
three_two_box <- create_three_two_box()
single_flag_box <- create_single_flag_box()
single_flag_box_2 <- create_single_flag_box_2()
three_flag_row_box <- create_three_flag_row_box()

### DEFINE ALL BOX CONFIGURATIONS ###
# Here I position the boxes relative to the overall map
box_configs <- list(
  list(plot = sahel_box, left = 0.27, bottom = 0.68, w = 245/2232, h = 205/1783),
  list(plot = lake_chad_box, left = 0.4, bottom = 0.55, w = 235/2232, h = 145/1783),
  list(plot = great_lakes_box, left = 0.52, bottom = 0.4, w = 177/2232, h = 177/1783),
  list(plot = two_row_box, left = 0.71, bottom = 0.22, w = 195/2232, h = 135/1783),
  list(plot = three_two_box, left = 0.76, bottom = 0.35, w = 245/2232, h = 138/1783),
  list(plot = single_flag_box, left = 0.70, bottom = 0.50, w = 95/2232, h = 65/1783),
  list(plot = single_flag_box_2, left = 0.58, bottom = 0.577, w = 95/2232, h = 65/1783),
  list(plot = three_flag_row_box, left = 0.515, bottom = 0.75, w = 205/2232, h = 84/1783)
)
```
### ADD ALL BOXES TO MAP
```{r}
for (config in box_configs) {
  base_map <- add_inset(base_map, config$plot, config$left, config$bottom, config$w, config$h)
}
```
## KEY
- This creates the grey legend arc plot data using the same method as the other arc plots.
### GREY ARC
```{r}
grey_legend_data <- tibble(
  segment = 1:3,
  label = c("A", "B", "C"),
  value = c("A", "B", "C"),
  color = "#D7D7D5",
  original_segment = 1:3
) |>
  arrange(desc(row_number())) |>
  mutate(
    segment = row_number(),
    original_start = pi - ((original_segment - 1) / 3 * pi),
    original_end = pi - (original_segment / 3 * pi),
    start_angle = (pi - ((segment - 1) / 3 * pi)) - pi/2,
    end_angle = (pi - (segment / 3 * pi)) - pi/2,
    mid_angle = (original_start + original_end) / 2,
    label_x = cos(mid_angle) * 0.77,
    label_y = sin(mid_angle) * 0.77
  )

# Plotting the grey legend arc plot
grey_legend_arc <- ggplot(grey_legend_data) +
  annotate("rect", xmin = -1.2, xmax = 1.2, ymin = -0.6, ymax = 1,
           fill = "white", color = NA) +
  geom_arc_bar(aes(x0 = 0, y0 = -0.3, r0 = 0.6, r = 0.9,
                   start = start_angle, end = end_angle, fill = color),
               color = "white", linewidth = 0.2) +
  scale_fill_identity() +
  geom_text(aes(x = label_x * 1.05, y = (label_y - 0.3) * 1.05, label = value),
            fontface = "bold", size = 2.75, lineheight = 0.9) +
  coord_fixed(xlim = c(-1.3, 1.3), ylim = c(-0.6, 1.0)) +
  theme_void() +
  theme(plot.background = element_rect(fill = "white", color = NA),
        plot.margin = margin(0, 0, 0, 0),
        text = element_text(size = rel(1)))
```
### TEXT
```{r}
white_rectangle <- ggplot() +
  annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1,
           fill = "white") +
  
  # Text - line 1
  annotate("text", x = 0.02, y = 1.00,
           label = "ACGRI pillars: IISS calculation based on multiple sources",
           hjust = 0, vjust = 1, size = 2.5) +
  
  # Text - line 2
  annotate("text", x = 0.02, y = 0.93,
           label = "for 2021 and January–April 2022 (scale: 0–100)",
           hjust = 0, vjust = 1, size = 2.5) +
  
  # Text to the right of arc - line 1
  annotate("text", x = 0.35, y = 0.80,
           label = "A: human impact, B: Incidence;",
           hjust = 0, vjust = 1, size = 2.5) +
  
  # Text to the right of arc - line 2
  annotate("text", x = 0.35, y = 0.73,
           label = "C: Geopolitical impact",
           hjust = 0, vjust = 1, size = 2.5) +
  
  # tText below arc
  annotate("text", x = 0.02, y = 0.59,
         label = "Number of fatalities, March 2021–April 2022",
         hjust = 0, vjust = 1, size = 2.5) +
  
  # Text left of bar
  annotate("text", x = 0.02, y = 0.49, label = "High", hjust = 0, vjust = 0.5, size = 2.5) +
  
  # Text right of bar
  annotate("text", x = 0.62, y = 0.49, label = "Low", hjust = 0, vjust = 0.5, size = 2.5) +
  
  # Text below bar
  annotate("text", x = 0.02, y = 0.41,
         label = "Involvement of third-party countries",
         hjust = 0, vjust = 1, size = 2.5) +
  
  # Peacekeeping text
  annotate("text", x = 0.02, y = 0.23, 
           label = "Peacekeeping, stabilisation and military-training operations",
           hjust = 0, vjust = 1, size = 2.5) +
  
  coord_fixed(ratio = 582/833, xlim = c(0, 1), ylim = c(0, 1)) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", color = "black", linewidth = 0.5),
    plot.margin = margin(0, 0, 0, 0)
  )
```
### POSITION AND ADD
```{r}
white_rect_w <- 833 / 2232
white_rect_h <- 582 / 1783
white_rect_left <- 50 / 2232
white_rect_bottom <- 80 / 1783

# Add white rectangle to map
base_map <- base_map +
  inset_element(
    white_rectangle,
    left = white_rect_left,
    right = white_rect_left + white_rect_w,
    bottom = white_rect_bottom,
    top = white_rect_bottom + white_rect_h
  )

# Add grey arc
grey_arc_w_map <- 215 / 2232
grey_arc_h_map <- 156 / 1783

# Position it inside the white rectangle
grey_arc_left <- white_rect_left + (20 / 2232)
grey_arc_bottom <- white_rect_bottom + (345 / 1783)

base_map <- base_map +
  inset_element(
    grey_legend_arc,
    left = grey_arc_left,
    right = grey_arc_left + grey_arc_w_map,
    bottom = grey_arc_bottom,
    top = grey_arc_bottom + grey_arc_h_map
  )
```
### HORIZONTAL BAR
```{r}
# Colours
color_scale_data <- tibble(
  x = seq(0, 48, length.out = 100),
  y = 1,
  fill_color = map_chr(x, value_to_color)
)

# Colour bar
color_scale_bar_only <- ggplot(color_scale_data, aes(x = x, y = y, fill = fill_color)) +
  geom_tile(width = 0.5, height = 1) +
  scale_fill_identity() +
  scale_x_reverse(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_fixed(ratio = 5, xlim = c(0, 48), ylim = c(0.5, 1.5)) +
  theme_void() +
  theme(
    plot.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(0, 0, 0, 0)
  )
```
### FLAGS COUNTRIUES
```{r}
create_involvement_flags <- function() {
  # Load flag images
  burundi_flag <- readPNG("flags_and_logos/burundi.png")
  eritrea_flag <- readPNG("flags_and_logos/eritrea.png")
  france_flag <- readPNG("flags_and_logos/france.png")
  italy_flag <- readPNG("flags_and_logos/italy.png")
  russia_flag <- readPNG("flags_and_logos/russia.png")
  rwanda_flag <- readPNG("flags_and_logos/rwanda.png")
  south_africa_flag <- readPNG("flags_and_logos/south_africa.png")
  turkey_flag <- readPNG("flags_and_logos/turkey.png")
  uk_flag <- readPNG("flags_and_logos/uk.png")
  us_flag <- readPNG("flags_and_logos/us.png")
  
  # Flag dimensions and spacing
  flag_width_px <- 54
  flag_height_px <- 37
  flag_width <- flag_width_px / 833
  flag_height <- flag_height_px / 582
  n_flags <- 10
  total_flags_width <- n_flags * flag_width
  spacing <- (1 - total_flags_width) / (n_flags + 1)
  
  # Helper function to calculate flag position
  flag_x <- function(i) spacing + (i * (flag_width + spacing))
  
  p <- ggplot() +
    annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1, fill = NA, color = NA) +
    
    # Burundi
    annotation_raster(burundi_flag,
                      xmin = flag_x(0), xmax = flag_x(0) + flag_width,
                      ymin = 0.6, ymax = 0.6 + flag_height) +
    annotate("text", x = flag_x(0) + flag_width/2, y = 0.58, 
             label = "Burundi", hjust = 0.5, vjust = 1, size = 2) +
    
    # Eritrea
    annotation_raster(eritrea_flag,
                      xmin = flag_x(1), xmax = flag_x(1) + flag_width,
                      ymin = 0.6, ymax = 0.6 + flag_height) +
    annotate("text", x = flag_x(1) + flag_width/2, y = 0.58, 
             label = "Eritrea", hjust = 0.5, vjust = 1, size = 2) +
    
    # France
    annotation_raster(france_flag,
                      xmin = flag_x(2), xmax = flag_x(2) + flag_width,
                      ymin = 0.6, ymax = 0.6 + flag_height) +
    annotate("text", x = flag_x(2) + flag_width/2, y = 0.58, 
             label = "France", hjust = 0.5, vjust = 1, size = 2) +
    
    # Italy
    annotation_raster(italy_flag,
                      xmin = flag_x(3), xmax = flag_x(3) + flag_width,
                      ymin = 0.6, ymax = 0.6 + flag_height) +
    annotate("text", x = flag_x(3) + flag_width/2, y = 0.58, 
             label = "Italy", hjust = 0.5, vjust = 1, size = 2) +
    
    # Russia
    annotation_raster(russia_flag,
                      xmin = flag_x(4), xmax = flag_x(4) + flag_width,
                      ymin = 0.6, ymax = 0.6 + flag_height) +
    annotate("text", x = flag_x(4) + flag_width/2, y = 0.58, 
             label = "Russia", hjust = 0.5, vjust = 1, size = 2) +
    
    # Rwanda
    annotation_raster(rwanda_flag,
                      xmin = flag_x(5), xmax = flag_x(5) + flag_width,
                      ymin = 0.6, ymax = 0.6 + flag_height) +
    annotate("text", x = flag_x(5) + flag_width/2, y = 0.58, 
             label = "Rwanda", hjust = 0.5, vjust = 1, size = 2) +
    
    # South Africa
    annotation_raster(south_africa_flag,
                      xmin = flag_x(6), xmax = flag_x(6) + flag_width,
                      ymin = 0.6, ymax = 0.6 + flag_height) +
    annotate("text", x = flag_x(6) + flag_width/2, y = 0.58, 
             label = "S.Africa", hjust = 0.5, vjust = 1, size = 2) +
    
    # Turkey
    annotation_raster(turkey_flag,
                      xmin = flag_x(7), xmax = flag_x(7) + flag_width,
                      ymin = 0.6, ymax = 0.6 + flag_height) +
    annotate("text", x = flag_x(7) + flag_width/2, y = 0.58, 
             label = "Turkey", hjust = 0.5, vjust = 1, size = 2) +
    
    # UK
    annotation_raster(uk_flag,
                      xmin = flag_x(8), xmax = flag_x(8) + flag_width,
                      ymin = 0.6, ymax = 0.6 + flag_height) +
    annotate("text", x = flag_x(8) + flag_width/2, y = 0.58, 
             label = "UK", hjust = 0.5, vjust = 1, size = 2) +
    
    # US
    annotation_raster(us_flag,
                      xmin = flag_x(9), xmax = flag_x(9) + flag_width,
                      ymin = 0.6, ymax = 0.6 + flag_height) +
    annotate("text", x = flag_x(9) + flag_width/2, y = 0.58, 
             label = "US", hjust = 0.5, vjust = 1, size = 2) +
    
    coord_fixed(ratio = 582/833, xlim = c(0, 1), ylim = c(0, 1)) +
    theme_void() +
    theme(
      plot.background = element_rect(fill = NA, color = NA),
      plot.margin = margin(0, 0, 0, 0)
    )
  
  return(p)
}
```
### ADD ALL PARTS TO THE WHITE BOX
```{r}
# Plot flags
flags_plot <- create_involvement_flags()

# Position the bar
bar_w_map <- 600 / 2232
bar_h_map <- 30 / 1783
bar_left <- white_rect_left + (0.02 * (833 / 2232)) + (5 / 2232)
bar_bottom <- white_rect_bottom + (260 / 1783)

# Add the bar
base_map <- base_map +
  inset_element(
    color_scale_bar_only,
    left = bar_left,
    right = bar_left + bar_w_map,
    bottom = bar_bottom,
    top = bar_bottom + bar_h_map
  )

# Add flags
flags_w_map <- 833 / 2232
flags_h_map <- 600 / 1783

base_map <- base_map +
  inset_element(
    flags_plot,
    left = white_rect_left,
    right = white_rect_left + flags_w_map,
    bottom = white_rect_bottom - (400 / 1783),
    top = white_rect_bottom + (50 / 1783) + flags_h_map
  )
```
### PEACEKEEPING LOGOS
```{r}
create_peacekeeping_logos <- function() {
  # Load logo images
  un_logo <- readPNG("flags_and_logos/un.png")
  eu_logo <- readPNG("flags_and_logos/eu.png")
  au_logo <- readPNG("flags_and_logos/au.png")
  g5_logo <- readPNG("flags_and_logos/sahel_g5.png")
  multi_logo <- readPNG("flags_and_logos/multi.png")
  sadcmm_logo <- readPNG("flags_and_logos/sadcmm.png")
  
  # Logo dimensions
  logo_width_px <- 54
  logo_height_px <- 37
  logo_width <- logo_width_px / 833
  logo_height <- logo_height_px / 582
  
  # Manual positioning
  spacing_left <- 0.02
  logo_gap <- 0.12
  logo_x <- function(i) spacing_left + (i * logo_gap)
  last_logo_x <- 0.75
  
  p <- ggplot() +
    annotate("rect", xmin = 0, xmax = 1, ymin = 0, ymax = 1, fill = NA, color = NA) +
    
    # UN
    annotation_raster(un_logo,
                      xmin = logo_x(0), xmax = logo_x(0) + logo_width,
                      ymin = 0.5, ymax = 0.5 + logo_height) +
    annotate("text", x = logo_x(0) + logo_width/2, y = 0.48, 
             label = "UN", hjust = 0.5, vjust = 1, size = 2, lineheight = 0.9) +
    
    # EU
    annotation_raster(eu_logo,
                      xmin = logo_x(1), xmax = logo_x(1) + logo_width,
                      ymin = 0.5, ymax = 0.5 + logo_height) +
    annotate("text", x = logo_x(1) + logo_width/2, y = 0.48, 
             label = "EU", hjust = 0.5, vjust = 1, size = 2, lineheight = 0.9) +
    
    # AU
    annotation_raster(au_logo,
                      xmin = logo_x(2), xmax = logo_x(2) + logo_width,
                      ymin = 0.5, ymax = 0.5 + logo_height) +
    annotate("text", x = logo_x(2) + logo_width/2, y = 0.48, 
             label = "African\nUnion", hjust = 0.5, vjust = 1, size = 2, lineheight = 0.9) +
    
    # G5 Sahel
    annotation_raster(g5_logo,
                      xmin = logo_x(3), xmax = logo_x(3) + logo_width,
                      ymin = 0.5, ymax = 0.5 + logo_height) +
    annotate("text", x = logo_x(3) + logo_width/2, y = 0.48, 
             label = "G5\nSahel", hjust = 0.5, vjust = 1, size = 2, lineheight = 0.9) +
    
    # MNJTF
    annotation_raster(multi_logo,
                      xmin = logo_x(4), xmax = logo_x(4) + logo_width,
                      ymin = 0.5, ymax = 0.5 + logo_height) +
    annotate("text", x = logo_x(4) + logo_width/2, y = 0.48, 
             label = "Multinational\nJoint Task\nForce", hjust = 0.5, vjust = 1, size = 2, lineheight = 0.9) +
    
    # SADCMM
    annotation_raster(sadcmm_logo,
                      xmin = last_logo_x, xmax = last_logo_x + logo_width,
                      ymin = 0.5, ymax = 0.5 + logo_height) +
    annotate("text", x = last_logo_x + logo_width/2, y = 0.48, 
             label = "Southern African\nDevelopment Community\nMission in Mozambique", hjust = 0.5, vjust = 1, size = 2, lineheight = 0.9) +
    
    coord_fixed(ratio = 582/833, xlim = c(0, 1), ylim = c(0, 1)) +
    theme_void() +
    theme(
      plot.background = element_rect(fill = NA, color = NA),
      plot.margin = margin(0, 0, 0, 0)
    )
  
  return(p)
}

# Create logos plot
logos_plot <- create_peacekeeping_logos()

logos_w_map <- 833 / 2232
logos_h_map <- 200 / 1783

base_map <- base_map +
  inset_element(
    logos_plot,
    left = white_rect_left,
    right = white_rect_left + logos_w_map,
    bottom = white_rect_bottom - (620 / 1783),
    top = white_rect_bottom + (590 / 1783) + logos_h_map
  )
```
## PLOT FINAL MAP!
```{r}
base_map
```

